---
title: "MC_paper2_phyloseq_import"
author: "Zeya Zhengyao Xue"
date: "January 14, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file follows the analysis in QIIME2, which imports mapping file, feature table,
taxonomy and tree as a 4-component phyloseq object. 

## Setting up packages and working directory 
```{r}
library(phyloseq);packageVersion("phyloseq")
library(ggplot2);packageVersion("ggplot2")
library(reshape2);packageVersion("reshape2")
library(cowplot);packageVersion("cowplot")
library(plyr)
library(dplyr)

path <- "G:/My Drive/UC_Davis/Marco_lab/milk_microbiota/Mock_community/MC_paper2_QIIME2/"
```

## Import QIIME2 output and make phyloseq object
```{r}
# Transpose the QIIME2 output in excel
# For the theoretical values, I can adjust its absolute reads to suit data set by
# changing values in the feature table
SeqTab <- read.table(file.path(path,"feature_table/4file-combined_table-dada2.txt"), header = TRUE, stringsAsFactors = FALSE)
colnames(SeqTab) <- gsub("X","",colnames(SeqTab))
row.names(SeqTab) <- SeqTab$OTUID
SeqTab <- SeqTab[,-1]
SeqTab <- as.matrix.data.frame(SeqTab)

# Read in metadata file
samdf <- read.csv(file.path(path,"mapping/MC_paper2_samdf.csv"))
rownames(samdf) <- samdf$SampleID

# Read in exported rooted tree 
tree <- read_tree(file.path(path, "tree.nwk"))

# Parse out taxonomic assignment in excel and remove confidence column
# keep only taxa included in the "feature-table-mc2.csv" file
TaxTab <- read.table(file.path(path,"taxonomy.tsv"), header = TRUE, sep = '\t', na.strings = NA)
rownames(TaxTab) <- TaxTab$FeatureID
TaxTab <- TaxTab[,-1]
TaxTab <- as.matrix.data.frame(TaxTab)

# merge as the phyloseq ready to use
ps <- phyloseq(otu_table(SeqTab, taxa_are_rows = FALSE), tax_table(TaxTab),sample_data(samdf), phy_tree(tree))
ps # 3019 taxa and 161 sample

# Filter out singleton
ps <- filter_taxa(ps, function (x) {sum(x > 0) > 1}, prune=TRUE)
ps # 720 taxa and 161 samples
write.csv(ps@otu_table, file.path(path,"feature_table/feature-table-mc2.csv"))
```

## Inspect Library Sizes
```
#add a colum for read counts/library size
samdf$LibrarySize <- sample_sums(ps)
#sort by the library size in increasing order
samdf <- samdf[order(samdf$LibrarySize),]
samdf$Index <- seq(nrow(samdf))
ggplot(samdf, aes(x=Index, y=LibrarySize, color=SampleOrCtrl))+geom_point()
```
As expected, some of the UHT_NC had pretty high read coverage.
Also, I decided to not do decontam with the MC data because I know already which
taxa are contamination and which taxa are expected.


## Filter ASVs based on taxonomy and abundance 
```{r}
# Show available ranks in the dataset
rank_names(ps)
# Create table, number of features for each phylum
table(tax_table(ps)[, "Phylum"], exclude = NULL)
# remove asv that was not assigned at phylum level
ps.Phy <- subset_taxa(ps, !is.na(Phylum) & !Phylum %in% c(""))
ps.Phy # 687 taxa and 161 samples  
# Explore abundance to decide filter threshold
## Compute prevalence of each feature, store as data.frame
prevdf <- apply(X = otu_table(ps.Phy),
                MARGIN = ifelse(taxa_are_rows(ps.Phy), yes = 1, no = 2),
                FUN = function(x)(sum(x>0)))
# Add taxonomy and total read counts to this data.frame
prevdf <- data.frame(Prevalence = prevdf,
                     TotalAbundance = taxa_sums(ps.Phy),
                     tax_table(ps.Phy))
# plot the abudance per phylum 
ggplot(prevdf, aes(TotalAbundance, Prevalence / nsamples(ps.Phy),color=Phylum)) +
  # Include a guess for parameter
  geom_hline(yintercept = 0.01,  linetype = 2) + geom_point(size = 2, alpha = 0.7) +
  scale_x_log10() +  xlab("Total Abundance") + ylab("Prevalence [Frac. Samples]") +
  facet_wrap(~Phylum) + theme(legend.position="none")
# Most bacteria are in the Actinobacteria, Bacteroidetes, Firmircutes and Proteobacteria
# But I don't actually have Bacteroidetes members...

# Remove very low abudance ASVs independent of sample prevalence
x = taxa_sums(ps.Phy)
keepTaxa = which((x / sum(x)) > 0.00005)
ps.Phy05 <- prune_taxa(names(keepTaxa), ps.Phy)
ps.Phy05 # 155 taxa and 161 samples
```

## Determine rarefaction level for Alpha and Beta analysis later on
```
set.seed(123)
# Define function to calculate alpha diversity and rarefaction levels
calculate_rarefaction_curves <- function(psdata, measures, depths) {
  estimate_rarified_richness <- function(psdata, measures, depth) {
    if(max(sample_sums(psdata)) < depth) return()
    psdata <- prune_samples(sample_sums(psdata) >= depth, psdata)
    rarified_psdata <- rarefy_even_depth(psdata, depth, verbose = FALSE)
    alpha_diversity <- estimate_richness(rarified_psdata, measures = measures)
    # as.matrix forces the use of melt.array, which includes the Sample names (rownames)
    molten_alpha_diversity <- melt(as.matrix(alpha_diversity), varnames = c("SampleID", 'measures'), value.name = 'Alpha_diversity')
    molten_alpha_diversity
  }
  names(depths) <- depths # this enables automatic addition of the Depth to the output by ldply
  rarefaction_curve_data <- ldply(depths, estimate_rarified_richness, psdata = psdata, measures = measures, .id = 'Depth', .progress = ifelse(interactive(), 'text', 'none'))
  #convert Depth from factor to numeric
  rarefaction_curve_data$Depth <- as.numeric(levels(rarefaction_curve_data$Depth))[rarefaction_curve_data$Depth]
  rarefaction_curve_data
}
rarefaction_curve_data <- calculate_rarefaction_curves(ps.Phy05, c("Observed", "Shannon","Simpson"), 
                                                       rep(c(1, 2500, 5000, 7500, 10000), each = 10))
rarefaction_curve_data$SampleID <- gsub("X","",rarefaction_curve_data$SampleID)
# summarize alpha diversity
rarefaction_curve_data_summary <- ddply(rarefaction_curve_data, c('Depth', 'SampleID', 'measures'), 
                                        summarise, 
                                        Alpha_diversity_mean = mean(Alpha_diversity), 
                                        Alpha_diversity_sd = sd(Alpha_diversity))
# Add sample data
rarefaction_curve_data_summary_verbose <- merge(rarefaction_curve_data_summary, 
                                                data.frame(sample_data(ps.Phy)), 
                                                by.x = 'SampleID', by.y = 'row.names')
rarefaction_curve_data_summary_verbose <- rarefaction_curve_data_summary_verbose[, -1]

# plot out rarefaction curve
ggplot(rarefaction_curve_data_summary_verbose, aes(x = Depth, y = Alpha_diversity_mean, 
                                                   ymin = Alpha_diversity_mean - Alpha_diversity_sd, 
                                                   ymax = Alpha_diversity_mean + Alpha_diversity_sd,
                                                   color = SampleID, group = SampleID)) +
  scale_fill_brewer(palette = "Set2") + 
  geom_line()+
  geom_pointrange(size=0.1)+
  theme(legend.position="none") + #  remove legend
  facet_wrap(facets = ~ measures, scales = "free")
# export the library size to a csv file
write.csv(sample_sums(ps.Phy05), file.path(path, "mapping/library_size_ps.Phy05.csv"))

# The diversity leveled out after 2500 seqs per sample, rarefy at 4000
# to keep the sample with the fewest reads that is no a NC (MC.50.2)
ps.Phy.rare <- rarefy_even_depth(ps.Phy05, sample.size = 4000, replace = TRUE, 
                                 rngseed = 123 , trimOTUs = TRUE, verbose = TRUE) 
ps.Phy.rare # 155 taxa and 155 samples
```

## Divide the ps object for different dataset 
### For taxonomy analysis based on "ps.Phy.REC"
```
# Clean up the taxonomy level
#make the deepest taxonomic identification in TaxTab as the recommened taxonomy (REC)
RECps <- function(x, ps) {
  TaxTab2 <- as.data.frame(x)
  
  list.g <- as.character(TaxTab2$Genus)
  list.f <- as.character(TaxTab2$Family)
  list.o <- as.character(TaxTab2$Order)
  list.c <- as.character(TaxTab2$Class)
  list.p <- as.character(TaxTab2$Phylum)
  list.k <- as.character(TaxTab2$Kingdom)
  list.REC <- character(length(list.g))
  
  for(i in 1:dim(TaxTab2)[1]){
    G = which(TaxTab2$Genus[i] == "")
    F = which(TaxTab2$Family[i] == "")
    O = which(TaxTab2$Order[i] == "")
    C = which(TaxTab2$Class[i] == "")
    P = which(TaxTab2$Phylum[i] == "")
    K = which(TaxTab2$Kingdom[i] == "")
    if(length(G) == 0){
      list.REC[i] <- list.g[i]
    } else if(length(F) == 0){
      list.REC[i] <- list.f[i]
    } else if(length(O) == 0){
      list.REC[i] <- list.o[i]
    } else if(length(C) == 0){
      list.REC[i] <- list.c[i]
    } else if(length(P) == 0){
      list.REC[i] <- list.p[i]
    } else if(length(K) == 0){
      list.REC[i] <- list.k[i]
    } else
      list.REC[i] <- "meow"
  }
  
  TaxTab2$REC <- list.REC
  TaxTab2$REC <- factor(TaxTab2$REC)
  merge_phyloseq(ps, TaxTab2 %>% as.matrix() %>% tax_table())
}

ps.Phy.glom <- tax_glom(ps.Phy05, taxrank = "Genus", NArm = FALSE)
ps.Phy.REC <- RECps(TaxTab, ps.Phy.glom)
ps.Phy.REC # 91 taxa and 161 samples


# Define function to get the average value of the feature table 
AveFeatTab <- function(ps){
  df <- psmelt(ps)
  df.cast <- dcast(df, REC ~ For_AVE, mean, value.var = "Abundance")
  df.cast[, -1] <- lapply( df.cast[ , -1], function(x) x/sum(x, na.rm=TRUE) )
  df.cast
}

# 1. DNA extraction lysis method using the Total kit; BCMC1
samdf_BCMC1_total_lyse <- read.csv(file.path(path, "mapping/MC_paper2_samdf_BCMC1_total_lyse.csv"))
rownames(samdf_BCMC1_total_lyse) <- samdf_BCMC1_total_lyse$SampleID
ps1 <- phyloseq(sample_data(samdf_BCMC1_total_lyse), otu_table(ps.Phy.REC),
                tax_table(ps.Phy.REC), phy_tree(ps.Phy.REC))
ps1 # 91 taxa and 23 samples
write.csv(AveFeatTab(ps1), file.path(path, "feature_table/FeatTab_BCMC1_total_lyseAVE.csv"))





# DNA extraction kit choice on 10 ml UHT milk; BCMC2


# PMA treatment

```


### For alpha and beta div based ono ps.Phy.rare
```
# 1.rare 
ps1.rare <- phyloseq(sample_data(samdf_mc2_BCMC1_total_lyse), otu_table(ps.Phy.rare),
                     tax_table(ps.Phy.rare), phy_tree(ps.Phy.rare))
ps1.rare
```



