---
title: "MC_paper2_phyloseq_import"
author: "Zeya Zhengyao Xue"
date: "January 14, 2019"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.path = "figs/")
```

This file follows the analysis in QIIME2, which imports mapping file, feature table,
taxonomy and tree as a 4-component phyloseq object. 

## Setting up packages and working directory 
```{r}
library(phyloseq);packageVersion("phyloseq")
library(vegan);packageVersion("vegan")
library(ggplot2);packageVersion("ggplot2")
library(reshape2);packageVersion("reshape2")
library(cowplot);packageVersion("cowplot")
library(plyr)
library(dplyr)

path <- "~/Google Drive File Stream/My Drive/UC_Davis/Marco_lab/milk_microbiota/Mock_community/MC_paper2_QIIME2/"
```

## Import QIIME2 output and make phyloseq object
2/5/19: add Phase 3 results and import as the same ps object
```{r}
# Transpose the QIIME2 output in excel
# For the theoretical values, I can adjust its absolute reads to suit data set by
# changing values in the feature table
SeqTab <- read.table(file.path(path, "feature_table/5file-combined_table-dada2-mc2.txt"), header = TRUE, stringsAsFactors = FALSE)
colnames(SeqTab) <- gsub("X","",colnames(SeqTab))
row.names(SeqTab) <- SeqTab$OTUID
SeqTab <- SeqTab[,-1]
SeqTab <- as.matrix.data.frame(SeqTab)

# Read in metadata file
samdf <- read.csv(file.path(path,"mapping/5file-MC_paper2_samdf.csv"))
rownames(samdf) <- samdf$SampleID

# Read in exported rooted tree 
tree <- read_tree(file.path(path,"5file-tree.nwk"))

# Parse out taxonomic assignment in excel and remove confidence column
# keep only taxa included in the "feature-table-mc2.csv" file
TaxTab <- read.table(file.path(path,"5file-taxonomy-mc2.tsv"), header = TRUE, sep = '\t', na.strings = NA)
rownames(TaxTab) <- TaxTab$FeatureID
TaxTab <- TaxTab[,-1]
TaxTab <- as.matrix.data.frame(TaxTab)

# merge as the phyloseq ready to use
ps <- phyloseq(otu_table(SeqTab, taxa_are_rows = TRUE), tax_table(TaxTab),sample_data(samdf), phy_tree(tree))
ps # 1052 taxa and 212 samples 

# Filter out singleton
ps <- filter_taxa(ps, function (x) {sum(x > 0) > 1}, prune=TRUE)
ps # 1049 taxa and 212 samples
write.csv(ps@otu_table, file.path(path,"feature_table/5file-feature-table-mc2.csv"))
```

## Inspect Library Sizes
```{r}
#add a colum for read counts/library size
samdf$LibrarySize <- sample_sums(ps)
#sort by the library size in increasing order
samdf <- samdf[order(samdf$LibrarySize),]
samdf$Index <- seq(nrow(samdf))
ggplot(samdf, aes(x=Index, y=LibrarySize, color=SampleOrCtrl))+geom_point()
```
As expected, some of the UHT_NC had pretty high read coverage.
Also, I decided to not do decontam with the MC data because I know already which
taxa are contamination and which taxa are expected.


## Filter ASVs based on taxonomy and abundance 
```{r}
# Show available ranks in the dataset
rank_names(ps)
# Create table, number of features for each phylum
table(tax_table(ps)[, "Phylum"], exclude = NULL)
# remove asv that was not assigned at phylum level
ps.Phy <- subset_taxa(ps, !is.na(Phylum) & !Phylum %in% c(""))
ps.Phy # 1013 taxa and 212 samples
# Explore abundance to decide filter threshold
## Compute prevalence of each feature, store as data.frame
prevdf <- apply(X = otu_table(ps.Phy),
                MARGIN = ifelse(taxa_are_rows(ps.Phy), yes = 1, no = 2),
                FUN = function(x)(sum(x>0)))
# Add taxonomy and total read counts to this data.frame
prevdf <- data.frame(Prevalence = prevdf,
                     TotalAbundance = taxa_sums(ps.Phy),
                     tax_table(ps.Phy))
# plot the abudance per phylum 
ggplot(prevdf, aes(TotalAbundance, Prevalence / nsamples(ps.Phy),color=Phylum)) +
  # Include a guess for parameter
  geom_hline(yintercept = 0.01,  linetype = 2) + geom_point(size = 2, alpha = 0.7) +
  scale_x_log10() +  xlab("Total Abundance") + ylab("Prevalence [Frac. Samples]") +
  facet_wrap(~Phylum) + theme(legend.position="none")
# Most bacteria are in the Actinobacteria, Bacteroidetes, Firmircutes and Proteobacteria
# But I don't actually have Bacteroidetes members...

# Remove very low abudance ASVs independent of sample prevalence
x = taxa_sums(ps.Phy)
keepTaxa = which((x / sum(x)) > 0.00005)
ps.Phy05 <- prune_taxa(names(keepTaxa), ps.Phy)
ps.Phy05 # 197 taxa and 212 samples
```

## Determine rarefaction level for Alpha and Beta analysis later on
```{r}
set.seed(123)
# Define function to calculate alpha diversity and rarefaction levels
calculate_rarefaction_curves <- function(psdata, measures, depths) {
  estimate_rarified_richness <- function(psdata, measures, depth) {
    if(max(sample_sums(psdata)) < depth) return()
    psdata <- prune_samples(sample_sums(psdata) >= depth, psdata)
    rarified_psdata <- rarefy_even_depth(psdata, depth, verbose = FALSE)
    alpha_diversity <- estimate_richness(rarified_psdata, measures = measures)
    # as.matrix forces the use of melt.array, which includes the Sample names (rownames)
    molten_alpha_diversity <- melt(as.matrix(alpha_diversity), varnames = c("SampleID", 'measures'), value.name = 'Alpha_diversity')
    molten_alpha_diversity
  }
  names(depths) <- depths # this enables automatic addition of the Depth to the output by ldply
  rarefaction_curve_data <- ldply(depths, estimate_rarified_richness, psdata = psdata, measures = measures, .id = 'Depth', .progress = ifelse(interactive(), 'text', 'none'))
  #convert Depth from factor to numeric
  rarefaction_curve_data$Depth <- as.numeric(levels(rarefaction_curve_data$Depth))[rarefaction_curve_data$Depth]
  rarefaction_curve_data
}
rarefaction_curve_data <- calculate_rarefaction_curves(ps.Phy05, c("Observed", "Shannon","Simpson"), 
                                                       rep(c(1, 2500, 5000, 7500, 10000), each = 10))
rarefaction_curve_data$SampleID <- gsub("X","",rarefaction_curve_data$SampleID)
# summarize alpha diversity
rarefaction_curve_data_summary <- ddply(rarefaction_curve_data, c('Depth', 'SampleID', 'measures'), 
                                        summarise, 
                                        Alpha_diversity_mean = mean(Alpha_diversity), 
                                        Alpha_diversity_sd = sd(Alpha_diversity))
# Add sample data
rarefaction_curve_data_summary_verbose <- merge(rarefaction_curve_data_summary, 
                                                data.frame(sample_data(ps.Phy)), 
                                                by.x = 'SampleID', by.y = 'row.names')
rarefaction_curve_data_summary_verbose <- rarefaction_curve_data_summary_verbose[, -1]

# plot out rarefaction curve
ggplot(rarefaction_curve_data_summary_verbose, aes(x = Depth, y = Alpha_diversity_mean, 
                                                   ymin = Alpha_diversity_mean - Alpha_diversity_sd, 
                                                   ymax = Alpha_diversity_mean + Alpha_diversity_sd,
                                                   color = SampleID, group = SampleID)) +
  scale_fill_brewer(palette = "Set2") + 
  geom_line()+
  geom_pointrange(size=0.1)+
  theme(legend.position="none") + #  remove legend
  facet_wrap(facets = ~ measures, scales = "free")
# export the library size to a csv file
write.csv(sample_sums(ps.Phy05), file.path(path, "mapping/library_size_ps.Phy05_5file.csv"))

# The diversity leveled out after 2500 seqs per sample, rarefy at 3990
# to keep the sample with the fewest reads that is no a NC (MC.50.2)
ps.Phy.rare <- rarefy_even_depth(ps.Phy05, sample.size = 3990, replace = TRUE, 
                                 rngseed = 123 , trimOTUs = TRUE, verbose = TRUE) 
ps.Phy.rare # 197 taxa and 203 samples
```

## Divide the ps object for different dataset 
### For taxonomy analysis based on "ps.Phy.REC"
```{r}
# Clean up the taxonomy level
#make the deepest taxonomic identification in TaxTab as the recommened taxonomy (REC)
RECps <- function(x, ps) {
  TaxTab2 <- as.data.frame(x)
  
  list.g <- as.character(TaxTab2$Genus)
  list.f <- as.character(TaxTab2$Family)
  list.o <- as.character(TaxTab2$Order)
  list.c <- as.character(TaxTab2$Class)
  list.p <- as.character(TaxTab2$Phylum)
  list.k <- as.character(TaxTab2$Kingdom)
  list.REC <- character(length(list.g))
  
  for(i in 1:dim(TaxTab2)[1]){
    G = which(TaxTab2$Genus[i] == "")
    F = which(TaxTab2$Family[i] == "")
    O = which(TaxTab2$Order[i] == "")
    C = which(TaxTab2$Class[i] == "")
    P = which(TaxTab2$Phylum[i] == "")
    K = which(TaxTab2$Kingdom[i] == "")
    if(length(G) == 0){
      list.REC[i] <- list.g[i]
    } else if(length(F) == 0){
      list.REC[i] <- list.f[i]
    } else if(length(O) == 0){
      list.REC[i] <- list.o[i]
    } else if(length(C) == 0){
      list.REC[i] <- list.c[i]
    } else if(length(P) == 0){
      list.REC[i] <- list.p[i]
    } else if(length(K) == 0){
      list.REC[i] <- list.k[i]
    } else
      list.REC[i] <- "meow"
  }
  
  TaxTab2$REC <- list.REC
  TaxTab2$REC <- factor(TaxTab2$REC)
  merge_phyloseq(ps, TaxTab2 %>% as.matrix() %>% tax_table())
}

ps.Phy.glom <- tax_glom(ps.Phy05, taxrank = "Genus", NArm = FALSE)
ps.Phy.REC <- RECps(TaxTab, ps.Phy.glom)
ps.Phy.REC # 100 taxa and 212 samples


# Define function to get the average and raw value of the feature table 
AveFeatTab <- function(ps){
  df <- psmelt(ps)
  df.cast <- dcast(df, REC ~ For_AVE, mean, value.var = "Abundance")
  df.cast[, -1] <- lapply( df.cast[ , -1], function(x) x/sum(x, na.rm=TRUE) )
  df.cast
}
FeatTab <- function(ps){
  df <- psmelt(ps)
  df.cast <- dcast(df, For_AVE + SampleID ~ REC, value.var = "Abundance")
  df.cast
}


# 1. DNA extraction lysis method using the Total kit; BCMC1
samdf_BCMC1_total_lyse <- read.csv(file.path(path, "mapping/MC_paper2_samdf_BCMC1_total_lyse.csv"))
rownames(samdf_BCMC1_total_lyse) <- samdf_BCMC1_total_lyse$SampleID
ps1 <- phyloseq(sample_data(samdf_BCMC1_total_lyse), otu_table(ps.Phy.REC),
                tax_table(ps.Phy.REC), phy_tree(ps.Phy.REC))
ps1 # 91 taxa and 23 samples
write.csv(AveFeatTab(ps1), file.path(path, "feature_table/FeatTab_BCMC1_total_lyseAVE.csv"))


# 2. DNA extraction kit with MoBio
samdf_BCMC1_MoBio <- read.csv(file.path(path, "mapping/MC_paper2_samdf_BCMC1_MoBio.csv"))
rownames(samdf_BCMC1_MoBio) <- samdf_BCMC1_MoBio$SampleID
ps2 <- phyloseq(sample_data(samdf_BCMC1_MoBio), otu_table(ps.Phy.REC),
                tax_table(ps.Phy.REC), phy_tree(ps.Phy.REC))
ps2 # 91 taxa and 6 samples
write.csv(AveFeatTab(ps2), file.path(path, "feature_table/FeatTab_BCMC1_MoBioAVE.csv"))

# 3. DNA extraction kit choice on 10 ml UHT milk; BCMC2
samdf_BCMC2_kit_10ml <- read.csv(file.path(path, "mapping/MC_paper2_samdf_BCMC2_kit_10ml.csv"))
rownames(samdf_BCMC2_kit_10ml) <- samdf_BCMC2_kit_10ml$SampleID
ps3 <- phyloseq(sample_data(samdf_BCMC2_kit_10ml), otu_table(ps.Phy.REC),
                tax_table(ps.Phy.REC), phy_tree(ps.Phy.REC))
ps3 # 91 taxa and 30 samples
write.csv(AveFeatTab(ps3), file.path(path, "feature_table/FeatTab_BCMC2_kit_10mlAVE.csv"))

# 4. PMA treatment
samdf_PMA <- read.csv(file.path(path, "mapping/MC_paper2_samdf_PMA.csv"))
rownames(samdf_PMA) <- samdf_PMA$SampleID
ps4 <- phyloseq(sample_data(samdf_PMA), otu_table(ps.Phy.REC),
                tax_table(ps.Phy.REC), phy_tree(ps.Phy.REC))
ps4 # 91 taxa and 15 samples
write.csv(AveFeatTab(ps4), file.path(path, "feature_table/FeatTab_PMA.csv"))

# 5. Sample storage
samdf_sample_storage <- read.csv(file.path(path, "mapping/MC_paper2_samdf_sample_storage.csv"))
rownames(samdf_sample_storage) <- samdf_sample_storage$SampleID
ps5 <- phyloseq(sample_data(samdf_sample_storage), otu_table(ps.Phy.REC),
                tax_table(ps.Phy.REC), phy_tree(ps.Phy.REC))
ps5 # 91 taxa and 24 samples
write.csv(AveFeatTab(ps5), file.path(path, "feature_table/FeatTab_sample_storage.csv"))

# 6. Sample Volume
samdf_total_vol <- read.csv(file.path(path, "mapping/MC_paper2_samdf_total_vol.csv"))
rownames(samdf_total_vol) <- samdf_total_vol$SampleID
ps6 <- phyloseq(sample_data(samdf_total_vol), otu_table(ps.Phy.REC),
                tax_table(ps.Phy.REC), phy_tree(ps.Phy.REC))
ps6 # 91 taxa and 18 samples
write.csv(AveFeatTab(ps6), file.path(path, "feature_table/FeatTab_total_vol.csv"))

# 7. phase 3 kit, cleanup and lysing method 
samdf_phase3 <- read.csv(file.path(path, "mapping/phase3_samdf.csv"))
rownames(samdf_phase3) <- samdf_phase3$SampleID
ps7 <- phyloseq(sample_data(samdf_phase3), otu_table(ps.Phy.REC),
                tax_table(ps.Phy.REC), phy_tree(ps.Phy.REC))
ps7 # 100 taxa and 51 samples
write.csv(AveFeatTab(ps7), file.path(path, ("feature_table/FeatTab_phase3.csv")))


# create ps list with all phyloseq objects
psList <- list(ps1, ps2, ps3, ps4, ps5, ps6, ps7)
names(psList) <- c(1:7)

# use the function to write feature table (not average) list and write out csv
FeatTab.list <- lapply(psList, FeatTab)
mapply(write.csv, FeatTab.list, file = paste0(path, "feature_table/FeatTab_", names(FeatTab.list), ".csv"))
```

#### Taxonomy plot
```{r}

```


### For alpha and beta div based on ps.Phy.rare
```{r}
ps1.rare <- rarefy_even_depth(ps1, sample.size = 3990, replace = TRUE, rngseed = 123 , trimOTUs = TRUE, verbose = TRUE) 
ps2.rare <- rarefy_even_depth(ps2, sample.size = 3990, replace = TRUE, rngseed = 123 , trimOTUs = TRUE, verbose = TRUE) 
ps3.rare <- rarefy_even_depth(ps3, sample.size = 3990, replace = TRUE, rngseed = 123 , trimOTUs = TRUE, verbose = TRUE) 
ps4.rare <- rarefy_even_depth(ps4, sample.size = 3990, replace = TRUE, rngseed = 123 , trimOTUs = TRUE, verbose = TRUE) 
ps5.rare <- rarefy_even_depth(ps5, sample.size = 3990, replace = TRUE, rngseed = 123 , trimOTUs = TRUE, verbose = TRUE) 
ps6.rare <- rarefy_even_depth(ps6, sample.size = 3990, replace = TRUE, rngseed = 123 , trimOTUs = TRUE, verbose = TRUE) 
ps7.rare <- rarefy_even_depth(ps7, sample.size = 3990, replace = TRUE, rngseed = 123 , trimOTUs = TRUE, verbose = TRUE) 

ps1.rare #  27 taxa and 23 samples
ps2.rare # 10 taxa and 6 samples
ps3.rare # 55 taxa and 29 samples
ps4.rare # 61 taxa and 15 samples
ps5.rare # 18 taxa and 24 samples
ps6.rare # 51 taxa and 18 samples
ps7.rare # 39 taxa and 51 samples
```


#### Beta diversity 
```{r}
# Define function to get distance matrix 
# I have to use distance() from vegan packages due to in compatibility 
# https://github.com/joey711/phyloseq/issues/918
DistTab <- function(ps, DistMethod){
  df <- psmelt(ps)
  df.cast <- dcast(df, For_AVE + SampleID ~ REC, value.var = "Abundance")
  df.cast <- df.cast[,-1]
  row.names(df.cast) <- df.cast[,1]
  df.cast <- df.cast[,-1]
  dist <- vegdist(df.cast, method = DistMethod) %>% as.matrix()
}

# create ps list with all phyloseq objects
psList.rare <- list(ps1.rare, ps2.rare, ps3.rare, ps4.rare, ps5.rare, ps6.rare, ps7.rare)
names(psList.rare) <- c(1:7)

# use the function to write bray distance list and write out as csv files
BrayList <- lapply(psList.rare, DistTab, DistMethod = "bray")
mapply(write.csv, BrayList, file = paste0(path, "distance/bray_", names(psList.rare), ".csv"))

# write out weighted list as well
WeightList <- lapply(psList.rare, UniFrac, weighted = TRUE)
WeightList <- lapply(WeightList, as.matrix)
mapply(write.csv, WeightList, file = paste0(path, "distance/weighted_", names(psList.rare), ".csv"))
```
```{r}
# Define function to make box plot and add p-values to the plot 




# Global test
compare_means(len ~ dose,  data = ToothGrowth, method = "anova")
compare_means(len ~ supp, data = ToothGrowth, 
              group.by = "dose")
# Default method = "kruskal.test" for multiple groups
ggboxplot(ToothGrowth, x = "dose", y = "len",
          color = "dose", palette = "jco")+stat_compare_means()
# Change method to anova
ggboxplot(ToothGrowth, x = "dose", y = "len",
          color = "dose", palette = "jco")+
  stat_compare_means(method = "anova")


# Import organized and subsetted (only extraction) distance csv file for making bar plots
Bray.ext <- read.csv(file.path(path, "distance/organized_distance_extraction_bray.csv"))

```


BBBBBB
USE CCA TRIPLOT TO SEE IF THERE'S ASSOCIATION BETWEEN METHOD AND TAXA

AAAAAA
Alpha could be more useful when using actual milk samples? 

Percent recovery of GBS DNA with meconium extractions
generated from various extraction kits. Data are mean ± S

