---
title: "Cheddar-only-phyloseq-analysis"
author: "Zeya Zhengyao Xue"
date: "May 14, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load needed packages
```{r}
library(phyloseq);packageVersion("phyloseq")
library(vegan)
library(ggplot2);packageVersion("ggplot2")
library(DESeq2);packageVersion("DESeq2")
library(reshape2)
library(cowplot)
library(plyr)
library(dplyr)
library(RColorBrewer)
```


## Subset to contain ONLY Cheddar cheese samples (exclude vintage Cheddar)
```{r}
path <- "G:/My Drive/UC_Davis/Marco_lab/milk_microbiota/Hilmar_weekly_samples/QIIME2/"

# Read in ps object that was NOT rarefied 
# (because after remove Lactococcus, read depth will be low)
# from the taxa DESeq analysis results  
# make sure that the cheese included are the same set with milk analysis
samdf.4hr <- read.csv(file.path(path, "mapping/Run1-5_samdf-mc2_4hrTL-NoExtremeMilk.csv"))
rownames(samdf.4hr) <- samdf.4hr$SampleID

ps.Cheddar.noLY <- readRDS(file.path(path,"cheddar/ps.chedY.noL.rds"))
ps.Cheddar.noLY
ps.Cheddar.noLY <- phyloseq(otu_table(ps.Cheddar.noLY, taxa_are_rows = FALSE),
                            tax_table(ps.Cheddar.noLY),
                            phy_tree(ps.Cheddar.noLY),
                            sample_data(samdf.4hr))
ps.Cheddar.noLY

ps.Cheddar.noLN <- readRDS(file.path(path,"cheddar/ps.chedN.noL.rds"))
ps.Cheddar.noLN




```

## 2-Lactococcus genus sequences removed
```{r}
alpha_measures = c("Observed", "Shannon") 
alpha.CheddarY <- estimate_richness(ps.Cheddar.noLY, split = TRUE, measures = alpha_measures)
alpha.CheddarN <- estimate_richness(ps.Cheddar.noLN, split = TRUE, measures = alpha_measures)
row.names(alpha.CheddarY) <- gsub("X","",row.names(alpha.CheddarY) )
row.names(alpha.CheddarN) <- gsub("X","",row.names(alpha.CheddarN) )
alpha.CheddarY$ID <- row.names(alpha.CheddarY)
alpha.CheddarN$ID <- row.names(alpha.CheddarN)
#add sample metadata information
alpha.CheddarY <- merge(alpha.CheddarY, data.frame(sample_data(ps.Cheddar.noLY)), by.x = 'ID', by.y = 'row.names')
alpha.CheddarN <- merge(alpha.CheddarN, data.frame(sample_data(ps.Cheddar.noLN)), by.x = 'ID', by.y = 'row.names')
#Keep only ID variable and variables
var <- c("Observed","Shannon","CheeseOutcome","SampleType","CheeseType") #should be modified for different problems
alpha.CheddarY1 <- alpha.CheddarY[var]
alpha.CheddarN1 <- alpha.CheddarN[var]
alpha.CheddarY1m <- melt(alpha.CheddarY1, id.vars = c("CheeseOutcome","SampleType","CheeseType"),
                      variable.name = "measure",
                      value.name = "value")
alpha.CheddarN1m <- melt(alpha.CheddarN1, id.vars = c("CheeseOutcome","SampleType","CheeseType"),
                      variable.name = "measure",
                      value.name = "value")
#need to reset the order of the factors because this is using molten data frame
alpha.CheddarY1m$SampleType <- factor(alpha.CheddarY1m$SampleType, levels = c("Cheese_0D","Cheese_30D","Cheese_90D","Cheese_120D"))
alpha.CheddarY1m$CheeseOutcome <- factor(alpha.CheddarY1m$CheeseOutcome, levels = c("no_slits","slits"))
alpha.CheddarN1m$SampleType <- factor(alpha.CheddarN1m$SampleType, levels = c("Cheese_0D","Cheese_30D","Cheese_90D","Cheese_120D"))
alpha.CheddarN1m$CheeseOutcome <- factor(alpha.CheddarN1m$CheeseOutcome, levels = c("no_slits","slits"))
```

### Alpha diversity by CheeseOutcome (PMA treated as main figure, untreated as supplemental figure)
```{r}
# Define function to plot Shannon and observed diversity for HTST_feed and HTST_milk
alphaHTSTfig <- function(x, samType, w, h, path.out){
  x <- subset(x, measure %in% c("Shannon", "Observed"))
  
  pdf(path.out, width = w, height = h)
  (ggplot(x, aes(x = SampleType, y = value, fill = CheeseOutcome))+ 
      geom_boxplot() +
      scale_fill_manual(values = c("no_slits" = "#a6cee3", "slits" = "#1f78b4")) +
      facet_wrap(facets = ~ measure, scales = "free_y") +
      theme_bw(base_size = 15)+
      scale_x_discrete(labels = c("Cheese_0D" = "0D", "Cheese_30D" = "30D", "Cheese_90D" = "90D", "Cheese_120D" = "120D"))+
      xlab(samType)) %>% print()
  dev.off()
}

# Generate figures for only HTST feed and HTST milk 
alphaHTSTfig(alpha.CheddarY1m, 
             samType = "PMA treated samples", w = 6.5, h = 2.4,
             path.out = file.path(path, "cheddar/alpha/alphaY.pdf"))
alphaHTSTfig(alpha.CheddarN1m, 
             samType = "PMA untreated samples", w = 6.5, h = 2.4,
             path.out = file.path(path, "cheddar/alpha/alphaN.pdf"))


# Define ANOVA for alpha diversity testing
AlphaANOVA <- function(x){
  fit.sha <- aov(value ~ CheeseOutcome, data = subset(x, measure %in% "Observed"))
  fit.sim <- aov(value ~ CheeseOutcome, data = subset(x, measure %in% "Shannon"))
  y <- summary(fit.sha)[[1]]
  y <- rbind(y, summary(fit.sim)[[1]])
  y
}
AlphaANOVA(subset(alpha.CheddarY1m, SampleType %in% "Cheese_0D"))
AlphaANOVA(subset(alpha.CheddarY1m, SampleType %in% "Cheese_30D"))
AlphaANOVA(subset(alpha.CheddarY1m, SampleType %in% "Cheese_90D"))
AlphaANOVA(subset(alpha.CheddarY1m, SampleType %in% "Cheese_120D"))
AlphaANOVA(subset(alpha.CheddarN1m, SampleType %in% "Cheese_0D"))
AlphaANOVA(subset(alpha.CheddarN1m, SampleType %in% "Cheese_30D"))
AlphaANOVA(subset(alpha.CheddarN1m, SampleType %in% "Cheese_90D"))
AlphaANOVA(subset(alpha.CheddarN1m, SampleType %in% "Cheese_120D"))
```

## NMDS analysis (PMA treated as main figure, untreated as supplemental figure)
I chose NMDS to be consistent with milk analysis
```{r echo=TRUE}
## Define function for NMDS ordinate and plot
NMDSfig <- function(ps, w, h, path.out){
  NMDS.wuf <- ordinate(ps, method = "NMDS", distance = "wunifrac")
  
  # Create cluster centroid locations for only the first 3 axes
  vector <- data.frame(NMDS.wuf["points"])
  # Add sample data to the vector
  vector$ID <- row.names(vector)
  vector <- merge(vector, 
                  sample_data(ps) %>% data.frame(),
                  by.x = "ID", by.y = "SampleID", 
                  all.x = TRUE, sort=FALSE)
  # calculate the mean/centroids of the vector
  center <- ddply(vector,
                  c("SampleType","CheeseOutcome"),  
                  summarize,
                  ave1 = mean(points.MDS1),
                  ave2 = mean(points.MDS2))
  # merge vector and center data frame together by SampleType and CheeseOutcome columns
  merge <- merge(vector, center, by = c("SampleType","CheeseOutcome"))
  # average per sample replicate
  merge.ave <- merge %>% group_by(SampleType, CheeseOutcome, Week) %>% 
    summarise(points.MDS1 = mean(points.MDS1),
              points.MDS2 = mean(points.MDS2),
              ave1 = mean(ave1),
              ave2 = mean(ave2))
  
  #split the data 
  dat_list <- split(merge.ave, merge.ave$SampleType)
  #use lapply to creat plot per SampleType
  fig_list <- lapply(dat_list, function(x){
    
    f1 <- ggplot(x, aes(points.MDS1, points.MDS2, color = CheeseOutcome)) + 
      geom_point(size = 1)+ 
      geom_point(aes(x = ave1, y = ave2), size = 3)+
      scale_color_manual(values=c("no_slits" = "#a6cee3", "slits" = "#1f78b4"))+
      geom_segment(aes(x = ave1, xend = points.MDS1,
                       y = ave2, yend = points.MDS2))+
      facet_wrap(~SampleType, scales = "fixed")+
      ylab("NMDS 2")+ 
      xlab("NMDS 1")+
      theme_bw() +
      theme(text = element_text(size = 15)) 
  })
  
  pdf(path.out, width = w, height = h)
  (plot_grid(fig_list[[1]] + theme(legend.position = "none"),
            fig_list[[3]] + theme(legend.position = "none"),
            fig_list[[4]] + theme(legend.position = "none"),
            fig_list[[2]] + theme(legend.position = "none"),
            nrow = 1, align="hv",  labels= c("A","B","C","D"))) %>% print()
  dev.off()
}

# beta plot per sample type
NMDSfig(ps.Cheddar.noLY, w = 10.6, h = 2.4, 
        path.out = file.path(path, "cheddar/beta/chedY.pdf"))
NMDSfig(ps.Cheddar.noLN, w = 10.6, h = 2.4, 
        path.out = file.path(path, "cheddar/beta/chedN.pdf"))


## Define function for Adonis on weighted unifrac distance
AdonisWuf <- function(ps){
  wuf.dist <- phyloseq::distance(ps, method = "wunifrac")  # dist matrix
  sampledf <- data.frame(sample_data(ps))  # data frame from the sample_data
  adonis(wuf.dist ~ CheeseOutcome, data = sampledf)
}

# Adonis test
AdonisWuf(subset_samples(ps.Cheddar.noLY, SampleType %in% "Cheese_0D"))
AdonisWuf(subset_samples(ps.Cheddar.noLY, SampleType %in% "Cheese_30D"))
AdonisWuf(subset_samples(ps.Cheddar.noLY, SampleType %in% "Cheese_90D"))
AdonisWuf(subset_samples(ps.Cheddar.noLY, SampleType %in% "Cheese_120D"))
AdonisWuf(subset_samples(ps.Cheddar.noLN, SampleType %in% "Cheese_0D"))
AdonisWuf(subset_samples(ps.Cheddar.noLN, SampleType %in% "Cheese_30D"))
AdonisWuf(subset_samples(ps.Cheddar.noLN, SampleType %in% "Cheese_90D"))
AdonisWuf(subset_samples(ps.Cheddar.noLN, SampleType %in% "Cheese_120D"))
```

## Get Lactobacillus sequences
```{r}
# PMA treated 
# Extract abundance matrix from the phyloseq object
OTU.Cheddar.noLY <- as(otu_table(ps.Cheddar.noLY), "matrix")
# transpose if necessary
if(taxa_are_rows(ps.Cheddar.noLY)){OTU.Cheddar.noLY <- t(OTU.Cheddar.noLY)}
# Coerce to data.frame
OTU.Cheddar.noLY <- as.data.frame(OTU.Cheddar.noLY)
write.csv(OTU.Cheddar.noLY, "decontam/phyloseq/cheese/OTU.Cheddar.noLY.csv")
# Get taxa table as well and then manually combine information from the 2 files together
TAX.Cheddar.noLY <- as(tax_table(ps.Cheddar.noLY), "matrix")
TAX.Cheddar.noLY <- as.data.frame(TAX.Cheddar.noLY)
write.csv(TAX.Cheddar.noLY, "decontam/phyloseq/cheese/TAX.Cheddar.noLY.csv")


# PMA untreated 
OTU.Cheddar.noLN = as(otu_table(ps.Cheddar.noLN), "matrix")
if(taxa_are_rows(ps.Cheddar.noLN)){OTU.Cheddar.noLN <- t(OTU.Cheddar.noLN)}
OTU.Cheddar.noLN = as.data.frame(OTU.Cheddar.noLN)
write.csv(OTU.Cheddar.noLN, "decontam/phyloseq/cheese/OTU.Cheddar.noLN.csv")
# Get taxa table as well and then manuallN combine information from the 2 files together
TAX.Cheddar.noLN <- as(tax_table(ps.Cheddar.noLN), "matrix")
TAX.Cheddar.noLN <- as.data.frame(TAX.Cheddar.noLN)
write.csv(TAX.Cheddar.noLN, "decontam/phyloseq/cheese/TAX.Cheddar.noLN.csv")
```



