---
title: "Milk-phyloseq-analysis"
author: "Zeya Zhengyao Xue"
date: "April 2, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setting up
```{r}
library(phyloseq)
library(ggplot2);packageVersion("ggplot2")
library(superheat)
library(reshape2)
library(cowplot)
library(plyr)
library(dplyr)
library(RColorBrewer)
library(vegan)


# Define project path
path <- "G:/My Drive/UC_Davis/Marco_lab/milk_microbiota/Hilmar_weekly_samples/QIIME2/"
```

## Subset to contain milk samples
```{r}
setwd("/Volumes/GoogleDrive/My Drive/UC_Davis/Marco_lab/milk_microbiota/Hilmar_weekly_samples/QIIME2")
ps05.Phy05.rare.tre.sample <- readRDS("decontam/phyloseq/ps05.Phy05.rare.tre.sample.rds")

ps.milk <- subset_samples(ps05.Phy05.rare.tre.sample, SampleType %in% c("HTST_milk","Raw_milk","HTST_feed"))
ps.milk #260 taxa and 349 samples
saveRDS(ps.milk,"decontam/phyloseq/milk/ps.milk.rds")

#PMA treated sample
ps.milkY <- subset_samples(ps.milk, PMA %in% "Y")
ps.milkY #260 taxa and 173 samples
saveRDS(ps.milkY, "decontam/phyloseq/milk/ps.milkY.rds")
#PMA untreated samples
ps.milkN <- subset_samples(ps.milk, PMA %in% "N")
ps.milkN #260 taxa and 176 samples
saveRDS(ps.milkN, "decontam/phyloseq/milk/ps.milkN.rds")
```

## Alpha diversity analysis
```{r}
alpha_measures = c("Observed", "Chao1", "Shannon", "Simpson") 
alpha.milk <- estimate_richness(ps.milk, split = TRUE, measures = alpha_measures) 
alpha.milkY <- estimate_richness(ps.milkY, split = TRUE, measures = alpha_measures)
alpha.milkN <- estimate_richness(ps.milkN, split = TRUE, measures = alpha_measures)
row.names(alpha.milk) <- gsub("X","",row.names(alpha.milk) )
row.names(alpha.milkY) <- gsub("X","",row.names(alpha.milkY) )
row.names(alpha.milkN) <- gsub("X","",row.names(alpha.milkN) )
alpha.milk$ID <- row.names(alpha.milk)
alpha.milkY$ID <- row.names(alpha.milkY)
alpha.milkN$ID <- row.names(alpha.milkN)
#add sample metadata information
alpha.milk <- merge(alpha.milk, data.frame(sample_data(ps.milk)), by.x = 'ID', by.y = 'row.names')
alpha.milkY <- merge(alpha.milkY, data.frame(sample_data(ps.milkY)), by.x = 'ID', by.y = 'row.names')
alpha.milkN <- merge(alpha.milkN, data.frame(sample_data(ps.milkN)), by.x = 'ID', by.y = 'row.names')
#Keep only ID variable and variables
var <- c("Observed","Chao1","Shannon","Simpson","Week","CheeseOutcome","SampleType") 
var2 <- c("Observed","Chao1","Shannon","Simpson","Week","CheeseOutcome","SampleType","PMA")#should be modified for different problems
alpha.milk <- alpha.milk[var2]
alpha.milkY1 <- alpha.milkY[var]
alpha.milkN1 <- alpha.milkN[var]
alpha.milk <- melt(alpha.milk, id.vars = c("Week","CheeseOutcome","SampleType","PMA"),
                   variable.name = "measure",
                   value.name = "value")
alpha.milkY1m <- melt(alpha.milkY1, id.vars = c("Week","CheeseOutcome","SampleType"),
                      variable.name = "measure",
                      value.name = "value")
alpha.milkN1m <- melt(alpha.milkN1, id.vars = c("Week","CheeseOutcome","SampleType"),
                      variable.name = "measure",
                      value.name = "value")
#need to reset the order of the factors because this is using molten data frame
alpha.milk$SampleType <- factor(alpha.milk$SampleType, levels = c("Raw_milk","HTST_feed","HTST_milk"))
alpha.milkY1m$SampleType <- factor(alpha.milkY1m$SampleType, levels = c("Raw_milk","HTST_feed","HTST_milk"))
alpha.milkY1m$Week <- factor(alpha.milkY1m$Week, levels = c("1","2","3","4","5","6","7","8","9","10","11"))
alpha.milkY1m$CheeseOutcome <- factor(alpha.milkY1m$CheeseOutcome, levels = c("no_slits","slits"))
alpha.milkN1m$SampleType <- factor(alpha.milkN1m$SampleType, levels = c("Raw_milk","HTST_feed","HTST_milk"))
alpha.milkN1m$Week <- factor(alpha.milkN1m$Week, levels = c("1","2","3","4","5","6","7","8","9","10","11"))
alpha.milkN1m$CheeseOutcome <- factor(alpha.milkN1m$CheeseOutcome, levels = c("no_slits","slits"))
```

After cleaning and organizing the data, we can make some plots for visualization:

### Plot to contain both PMA treated and untreated 
```{r}
AlphaMilkType <- function(x, w, h, path.out){
  x <- subset(x, measure %in% c("Observed", "Shannon"))
  
  pdf(path.out, w, h)
  (ggplot(x, aes(x=SampleType, y=value))+ 
  geom_boxplot() +
  facet_wrap(facets = ~ measure, scales = "free_y") + #strip.position="right"
  theme_bw(base_size = 15)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  xlab("")) %>% print()
  dev.off()
}

AlphaMilkType(alpha.milk, w = 7, h = 3.7, 
              path.out = file.path(path,"milk/alpha/milktype.pdf"))
AlphaMilkType(alpha.milkY1m, w = 6.5, h = 3.8,
              path.out = file.path(path,"milk/alpha/milktypeY.pdf"))
AlphaMilkType(alpha.milkN1m, w = 6.5, h = 3.8,
              path.out = file.path(path,"milk/alpha/milktypeN.pdf"))

attach(subset(alpha.milkY1m, measure == "Observed"))
pairwise.t.test(value, SampleType, p.adjust.method = "none")[[3]]
detach()
attach(subset(alpha.milkY1m, measure == "Shannon"))
pairwise.t.test(value, SampleType, p.adjust.method = "none")[[3]]
detach()

attach(subset(alpha.milkN1m, measure == "Observed"))
pairwise.t.test(value, SampleType, p.adjust.method = "none")[[3]]
detach()
attach(subset(alpha.milkN1m, measure == "Shannon"))
pairwise.t.test(value, SampleType, p.adjust.method = "none")[[3]]
detach()
```


### 1. Plot by week (PMA treated as main figure, untreated as supplemental figure)
```{r}
alpha.wkY <- ggplot(alpha.milkY1m, aes(x=Week, y=value, group=SampleType, color=SampleType))+
  geom_point(size=0.5) + 
  geom_smooth() +
  scale_color_manual(values=c("Raw_milk"="black", "HTST_feed"="red", "HTST_milk"="blue")) +
  facet_wrap(facets = ~ measure, scales = "free_y") + #strip.position="right"
  theme_bw()+
  ylab("Alpha diversity value")+
  xlab("PMA treated samples by collection week")
alpha.wkY

alpha.wkN <- ggplot(alpha.milkN1m, aes(x=Week, y=value, group=SampleType, color=SampleType))+
  geom_point(size=0.5) + 
  geom_smooth() +
  scale_color_manual(values=c("Raw_milk"="black", "HTST_feed"="red", "HTST_milk"="blue")) +
  facet_wrap(facets = ~ measure, scales = "free_y") + #strip.position="right"
  theme_bw()+
  ylab("Alpha diversity value")+
  xlab("PMA untreated samples by collection week")
alpha.wkN
#save 6*4
```

### 2. Plot by CheeseOutcome (PMA treated as main figure, untreated as supplemental figure)
```{r}
alpha.coY <- ggplot(alpha.milkY1m, aes(x=CheeseOutcome, y=value, color=SampleType))+ 
  geom_boxplot() +
  scale_color_manual(values=c("Raw_milk"="black", "HTST_feed"="red", "HTST_milk"="blue")) +
  facet_wrap(facets = ~ measure, scales = "free_y") + #strip.position="right"
  theme_bw()+
  ylab("Alpha diversity value")+
  xlab("PMA treated samples by cheese outcome")
alpha.coY

alpha.coN <- ggplot(alpha.milkN1m, aes(x=CheeseOutcome, y=value, color=SampleType))+ 
  geom_boxplot() +
  scale_color_manual(values=c("Raw_milk"="black", "HTST_feed"="red", "HTST_milk"="blue")) +
  facet_wrap(facets = ~ measure, scales = "free_y") + #strip.position="right"
  theme_bw()+
  ylab("Alpha diversity value")+
  xlab("PMA untreated samples by cheese outcome")
alpha.coN

alpha.coY1 <- ggplot(alpha.milkY1m, aes(x=SampleType, y=value, fill=CheeseOutcome))+ 
  geom_boxplot() +
  scale_fill_manual(values=c("no_slits"="#a6cee3", "slits"="#1f78b4")) +
  facet_wrap(facets = ~ measure, scales = "free_y") + #strip.position="right"
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  ylab("Alpha diversity value")+
  xlab("PMA treated samples by milk type")
alpha.coY1

alpha.coN1 <- ggplot(alpha.milkN1m, aes(x=SampleType, y=value, fill=CheeseOutcome))+ 
  geom_boxplot() +
  scale_fill_manual(values=c("no_slits"="#a6cee3", "slits"="#1f78b4")) +
  facet_wrap(facets = ~ measure, scales = "free_y") + #strip.position="right"
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  ylab("Alpha diversity value")+
  xlab("PMA untreated samples by milk type")
alpha.coN1

# Define function to plot Shannon and observed diversity for HTST_feed and HTST_milk
alphaHTSTfig <- function(x, samType, w, h, path.out){
  x <- subset(x, SampleType %in% c("HTST_feed", "HTST_milk"))
  x <- subset(x, measure %in% c("Shannon", "Observed")) 
         
  pdf(path.out, width = w, height = h)
  (ggplot(x, aes(x = SampleType, y = value, fill = CheeseOutcome))+ 
  geom_boxplot() +
  scale_fill_manual(values = c("no_slits" = "#a6cee3", "slits" = "#1f78b4")) +
  facet_wrap(facets = ~ measure, scales = "free_y") +
  theme_bw(base_size = 15)+
  xlab(samType)) %>% print()
  dev.off()
}

# Generate figures for only HTST feed and HTST milk 
alphaHTSTfig(alpha.milkY1m, 
             samType = "PMA treated samples", w = 8.5, h = 3.3,
             path.out = file.path(path, "milk/alpha/alphaY.pdf"))
alphaHTSTfig(alpha.milkN1m, 
             samType = "PMA untreated samples", w = 8.5, h = 3.3,
             path.out = file.path(path, "milk/alpha/alphaN.pdf"))
# ANOVA test on the diversity
AlphaANOVA(subset(alpha.milkY1m, SampleType %in% "HTST_feed"))
AlphaANOVA(subset(alpha.milkY1m, SampleType %in% "HTST_milk"))
AlphaANOVA(subset(alpha.milkN1m, SampleType %in% "HTST_feed"))
AlphaANOVA(subset(alpha.milkN1m, SampleType %in% "HTST_milk"))

```

## Beta diversity analysis
As of May 15th, only certain diatance matrix work due to conflict in vegan and phyloseq versions.
```{r}
#set the order of factors
sample_data(ps.milkY)$SampleType <- factor(sample_data(ps.milkY)$SampleType, levels = c("Raw_milk","HTST_feed","HTST_milk"))
sample_data(ps.milkN)$SampleType <- factor(sample_data(ps.milkN)$SampleType, levels = c("Raw_milk","HTST_feed","HTST_milk"))
```

### NMDS (PMA treated as main figure, untreated as supplemental figure)
```{r eval=FALSE, include=FALSE, results="hide"}
set.seed(123)
#bray PMA
NMDS.bray.milkY <- ordinate(ps.milkY, method = "NMDS", distance = "bray") #stress < 0.2

NMDS.bray.milkYp <- plot_ordination(ps.milkY, NMDS.bray.milkY,
                                    type = "sample",
                                    color = "CheeseOutcome") + 
  scale_color_manual(values=c("no_slits"="#a6cee3", "slits"="#1f78b4"))+ 
  stat_ellipse() +
  geom_point() + 
  facet_wrap(~SampleType)+
  ggtitle("PMA treated samples (NMDS of Bray-Curtis dissimilarity)")+
  theme_bw()#+
  #theme(text = element_text(size = 16),plot.title = element_text(size = 18, hjust = 0.5)) #to center the title of the plot
NMDS.bray.milkYp #I can save 12x4 to generate square output 
```

Similar commends were used for making plots with other clustering and distance method, see original Rmarkdown file if interested in raw codes.
```{r eval=FALSE, include=FALSE, results="hide"}
#bray NO PMA
NMDS.bray.milkN <- ordinate(ps.milkN, method = "NMDS", distance = "bray") #stress < 0.2

NMDS.bray.milkNp <- plot_ordination(ps.milkN, NMDS.bray.milkN,
                                    type = "sample",
                                    color = "CheeseOutcome") + 
  scale_color_manual(values=c("no_slits"="#a6cee3", "slits"="#1f78b4"))+ 
  stat_ellipse() +
  geom_point() + 
  facet_wrap(~SampleType)+
  ggtitle("PMA untreated samples (NMDS of Bray-Curtis dissimilarity)")+
  theme_bw()
NMDS.bray.milkNp
```

```{r eval=FALSE, include=FALSE, results="hide"}
#weighted unifrac PMA
NMDS.wuf.milkY <- ordinate(ps.milkY, method = "NMDS", distance = "wunifrac") #stress < 0.2
NMDS.wuf.milkYp <- plot_ordination(ps.milkY, NMDS.wuf.milkY,
                                   type = "sample",
                                   color = "CheeseOutcome") + 
  scale_color_manual(values=c("no_slits"="#a6cee3", "slits"="#1f78b4"))+ 
  stat_ellipse() +
  geom_point() + 
  facet_wrap(~SampleType)+
  ggtitle("PMA treated samples (NMDS of weighted unifrac distance)")+
  theme_bw()
NMDS.wuf.milkYp
```

```{r eval=FALSE, include=FALSE, results="hide"}
#weighted unifrac NO PMA
NMDS.wuf.milkN <- ordinate(ps.milkN, method = "NMDS", distance = "wunifrac") #stress < 0.2
NMDS.wuf.milkNp <- plot_ordination(ps.milkN, NMDS.wuf.milkN,
                                   type = "sample",
                                   color = "CheeseOutcome") + 
  scale_color_manual(values=c("no_slits"="#a6cee3", "slits"="#1f78b4"))+ 
  stat_ellipse() +
  geom_point() + 
  facet_wrap(~SampleType)+
  ggtitle("PMA untreated samples (NMDS of weighted unifrac distance)")+
  theme_bw()
NMDS.wuf.milkNp
```

### PCoA analysis (PMA treated as main figure, untreated as supplemental figure)
```{r eval=FALSE, include=FALSE}
#Bray-Curtis PMA
PCoA.bray.milkY <- ordinate(ps.milkY, method = "PCoA", distance = "bray")
PCoA.bray.milkYp <- plot_ordination(ps.milkY, PCoA.bray.milkY,
                                   type = "sample",
                                   color = "CheeseOutcome") + 
  scale_color_manual(values=c("no_slits"="#a6cee3", "slits"="#1f78b4"))+ 
  stat_ellipse() +
  geom_point() + 
  facet_wrap(~SampleType)+
  ggtitle("PMA treated samples (PCoA of Bray-Curtis dissimilarity)")+
  theme_bw()
PCoA.bray.milkYp

#Bray-Curtis NO PMA
PCoA.bray.milkN <- ordinate(ps.milkN, method = "PCoA", distance = "bray")
PCoA.bray.milkNp <- plot_ordination(ps.milkN, PCoA.bray.milkN,
                                   type = "sample",
                                   color = "CheeseOutcome") + 
  scale_color_manual(values=c("no_slits"="#a6cee3", "slits"="#1f78b4"))+ 
  stat_ellipse() +
  geom_point() + 
  facet_wrap(~SampleType)+
  ggtitle("PMA untreated samples (PCoA of Bray-Curtis dissimilarity)")+
  theme_bw()
PCoA.bray.milkNp

#weighted unifrac PMA
PCoA.wuf.milkY <- ordinate(ps.milkY, method = "PCoA", distance = "wunifrac")
PCoA.wuf.milkYp <- plot_ordination(ps.milkY, PCoA.bray.milkY,
                                   type = "sample",
                                   color = "CheeseOutcome") + 
  scale_color_manual(values=c("no_slits"="#a6cee3", "slits"="#1f78b4"))+ 
  stat_ellipse() +
  geom_point() + 
  facet_wrap(~SampleType)+
  ggtitle("PMA treated samples (PCoA of weighted unifrac distance")+
  theme_bw()
PCoA.wuf.milkYp

#weighted unifrac NO PMA
PCoA.wuf.milkN <- ordinate(ps.milkN, method = "PCoA", distance = "wunifrac")
PCoA.wuf.milkNp <- plot_ordination(ps.milkN, PCoA.bray.milkN,
                                   type = "sample",
                                   color = "CheeseOutcome") + 
  scale_color_manual(values=c("no_slits"="#a6cee3", "slits"="#1f78b4"))+ 
  stat_ellipse() +
  geom_point() + 
  facet_wrap(~SampleType)+
  ggtitle("PMA untreated samples (PCoA of weighted unifrac distance")+
  theme_bw()
PCoA.wuf.milkNp
```

## Beta diverity analysis 
### 1.1 CheeseOutcome with all milk
```{r fig.height=3, fig.width=13}
#set the order of factors
sample_data(ps.milkY)$SampleType <- factor(sample_data(ps.milkY)$SampleType, levels = c("Raw_milk","HTST_feed","HTST_milk"))
sample_data(ps.milkN)$SampleType <- factor(sample_data(ps.milkN)$SampleType, levels = c("Raw_milk","HTST_feed","HTST_milk"))

## PCoA 
# PMA treated
PCoA.wuf.milkY <- ordinate(ps.milkY, method = "PCoA", distance = "wunifrac")
# Create cluster centroid locations for only the first 3 axes
PCoA.wuf.milkY_vector <- data.frame(PCoA.wuf.milkY["vectors"])[,1:3]
# Add sample data to the vector
PCoA.wuf.milkY_vector$ID <- row.names(PCoA.wuf.milkY_vector)
PCoA.wuf.milkY_vector <- merge(PCoA.wuf.milkY_vector, 
                                sample_data(ps.milkY) %>% data.frame(),
                                by.x = "ID", by.y = "SampleID", 
                                all.x = TRUE, sort=FALSE)
#calculate the mean/centroids of the vector
PCoA.wuf.milkY_center <- ddply(PCoA.wuf.milkY_vector,
                                c("SampleType","CheeseOutcome"),
                                summarize,
                                ave1 = mean(vectors.Axis.1),
                                ave2 = mean(vectors.Axis.2),
                                ave3 = mean(vectors.Axis.3))
#merge vector and center data frame together by SampleType and CheeseOutcome columns
PCoA.wuf.milkY_merge <- merge(PCoA.wuf.milkY_vector,
                               PCoA.wuf.milkY_center,
                               by = c("SampleType","CheeseOutcome"))

#split the data 
dat_list <- split(PCoA.wuf.milkY_merge, PCoA.wuf.milkY_merge$SampleType)
#use lapply to creat plot for the 3 milk SampleType
fig_list <- lapply(dat_list, function(x){
  
  f1 <- ggplot(x, aes(vectors.Axis.1, vectors.Axis.2, color = CheeseOutcome)) + 
    geom_point(size = 1)+ 
    geom_point(aes(x = ave1, y = ave2), size = 3)+
    scale_color_manual(values=c("no_slits" = "#a6cee3", "slits" = "#1f78b4"))+
    geom_segment(aes(x = ave1, xend = vectors.Axis.1,
                     y = ave2, yend = vectors.Axis.2))+
    facet_wrap(~SampleType, scales = "fixed")+
    ylab("PCoA 2")+ 
    xlab("PCoA 1")+
    theme_bw()
})
#make the figure
legend <- get_legend(fig_list[[1]])
PCoA.wuf.milkYp <- plot_grid(fig_list[[1]] + theme(legend.position = "none"),
                              fig_list[[2]] + theme(legend.position = "none"),
                              fig_list[[3]] + theme(legend.position = "none"),
                              legend, nrow = 1, align="hv",  labels= c("A","B","C"))
PCoA.wuf.milkYp


# PMA untreated
PCoA.wuf.milkN <- ordinate(ps.milkN, method = "PCoA", distance = "wunifrac")
# Create cluster centroid locations for only the first 3 axes
PCoA.wuf.milkN_vector <- data.frame(PCoA.wuf.milkN["vectors"])[,1:3]
# Add sample data to the vector
PCoA.wuf.milkN_vector$ID <- row.names(PCoA.wuf.milkN_vector)
PCoA.wuf.milkN_vector <- merge(PCoA.wuf.milkN_vector, 
                                sample_data(ps.milkN) %>% data.frame(),
                                by.x = "ID", by.y = "SampleID", 
                                all.x = TRUE, sort=FALSE)
#calculate the mean/centroids of the vector
PCoA.wuf.milkN_center <- ddply(PCoA.wuf.milkN_vector,
                                c("SampleType","CheeseOutcome"),
                                summarize,
                                ave1 = mean(vectors.Axis.1),
                                ave2 = mean(vectors.Axis.2),
                                ave3 = mean(vectors.Axis.3))
#merge vector and center data frame together by SampleType and CheeseOutcome columns
PCoA.wuf.milkN_merge <- merge(PCoA.wuf.milkN_vector,
                               PCoA.wuf.milkN_center,
                               by = c("SampleType","CheeseOutcome"))

#split the data 
dat_list <- split(PCoA.wuf.milkN_merge, PCoA.wuf.milkN_merge$SampleType)
#use lapply to creat plot for the 3 milk SampleType
fig_list <- lapply(dat_list, function(x){
  
  f1 <- ggplot(x, aes(vectors.Axis.1, vectors.Axis.2, color = CheeseOutcome)) + 
    geom_point(size = 1)+ 
    geom_point(aes(x = ave1, y = ave2), size = 3)+
    scale_color_manual(values=c("no_slits" = "#a6cee3", "slits" = "#1f78b4"))+
    geom_segment(aes(x = ave1, xend = vectors.Axis.1,
                     y = ave2, yend = vectors.Axis.2))+
    facet_wrap(~SampleType, scales = "fixed")+
    ylab("PCoA 2")+ 
    xlab("PCoA 1")+
    theme_bw()
})
#make the figure
legend <- get_legend(fig_list[[1]])
PCoA.wuf.milkNp <- plot_grid(fig_list[[1]] + theme(legend.position = "none"),
                              fig_list[[2]] + theme(legend.position = "none"),
                              fig_list[[3]] + theme(legend.position = "none"),
                              legend, nrow = 1, align="hv",  labels= c("A","B","C"))
PCoA.wuf.milkNp


## NMDS 
# PMA treated
NMDS.wuf.milkY <- ordinate(ps.milkY, method = "NMDS", distance = "wunifrac")
# Create cluster centroid locations for only the first 3 axes
NMDS.wuf.milkY_vector <- data.frame(NMDS.wuf.milkY["points"])
# Add sample data to the vector
NMDS.wuf.milkY_vector$ID <- row.names(NMDS.wuf.milkY_vector)
NMDS.wuf.milkY_vector <- merge(NMDS.wuf.milkY_vector, 
                                sample_data(ps.milkY) %>% data.frame(),
                                by.x = "ID", by.y = "SampleID", 
                                all.x = TRUE, sort=FALSE)
#calculate the mean/centroids of the vector
NMDS.wuf.milkY_center <- ddply(NMDS.wuf.milkY_vector,
                                c("SampleType","CheeseOutcome"),
                                summarize,
                                ave1 = mean(points.MDS1),
                                ave2 = mean(points.MDS2))
#merge vector and center data frame together by SampleType and CheeseOutcome columns
NMDS.wuf.milkY_merge <- merge(NMDS.wuf.milkY_vector,
                               NMDS.wuf.milkY_center,
                               by = c("SampleType","CheeseOutcome"))

#split the data 
dat_list <- split(NMDS.wuf.milkY_merge, NMDS.wuf.milkY_merge$SampleType)
#use lapply to creat plot for the 3 milk SampleType
fig_list <- lapply(dat_list, function(x){
  
  f1 <- ggplot(x, aes(points.MDS1, points.MDS2, color = CheeseOutcome)) + 
    geom_point(size = 1)+ 
    geom_point(aes(x = ave1, y = ave2), size = 3)+
    scale_color_manual(values=c("no_slits" = "#a6cee3", "slits" = "#1f78b4"))+
    geom_segment(aes(x = ave1, xend = points.MDS1,
                     y = ave2, yend = points.MDS2))+
    facet_wrap(~SampleType, scales = "fixed")+
    ylab("NMDS 2")+ 
    xlab("NMDS 1")+
    theme_bw()
})
#make the figure
legend <- get_legend(fig_list[[1]])
NMDS.wuf.milkYp <- plot_grid(fig_list[[1]] + theme(legend.position = "none"),
                              fig_list[[2]] + theme(legend.position = "none"),
                              fig_list[[3]] + theme(legend.position = "none"),
                              legend, nrow = 1, align="hv",  labels= c("A","B","C"))
NMDS.wuf.milkYp

# PMA untreated
NMDS.wuf.milkN <- ordinate(ps.milkN, method = "NMDS", distance = "wunifrac")
# Create cluster centroid locations for only the first 3 axes
NMDS.wuf.milkN_vector <- data.frame(NMDS.wuf.milkN["points"])
# Add sample data to the vector
NMDS.wuf.milkN_vector$ID <- row.names(NMDS.wuf.milkN_vector)
NMDS.wuf.milkN_vector <- merge(NMDS.wuf.milkN_vector, 
                                sample_data(ps.milkN) %>% data.frame(),
                                by.x = "ID", by.y = "SampleID", 
                                all.x = TRUE, sort=FALSE)
#calculate the mean/centroids of the vector
NMDS.wuf.milkN_center <- ddply(NMDS.wuf.milkN_vector,
                                c("SampleType","CheeseOutcome"),
                                summarize,
                                ave1 = mean(points.MDS1),
                                ave2 = mean(points.MDS2))
#merge vector and center data frame together by SampleType and CheeseOutcome columns
NMDS.wuf.milkN_merge <- merge(NMDS.wuf.milkN_vector,
                               NMDS.wuf.milkN_center,
                               by = c("SampleType","CheeseOutcome"))

#split the data 
dat_list <- split(NMDS.wuf.milkN_merge, NMDS.wuf.milkN_merge$SampleType)
#use lapply to creat plot for the 3 milk SampleType
fig_list <- lapply(dat_list, function(x){
  
  f1 <- ggplot(x, aes(points.MDS1, points.MDS2, color = CheeseOutcome)) + 
    geom_point(size = 1)+ 
    geom_point(aes(x = ave1, y = ave2), size = 3)+
    scale_color_manual(values=c("no_slits" = "#a6cee3", "slits" = "#1f78b4"))+
    geom_segment(aes(x = ave1, xend = points.MDS1,
                     y = ave2, yend = points.MDS2))+
    facet_wrap(~SampleType, scales = "fixed")+
    ylab("NMDS 2")+ 
    xlab("NMDS 1")+
    theme_bw()
})
#make the figure
legend <- get_legend(fig_list[[1]])
NMDS.wuf.milkNp <- plot_grid(fig_list[[1]] + theme(legend.position = "none"),
                              fig_list[[2]] + theme(legend.position = "none"),
                              fig_list[[3]] + theme(legend.position = "none"),
                              legend, nrow = 1, align="hv",  labels= c("A","B","C"))
NMDS.wuf.milkNp
```

### 1.2 MilkType with all milk
```{r}
sample_data(ps.milkY)$SampleType <- factor(sample_data(ps.milkY)$SampleType, levels = c("Raw_milk","HTST_feed","HTST_milk"))
sample_data(ps.milkN)$SampleType <- factor(sample_data(ps.milkN)$SampleType, levels = c("Raw_milk","HTST_feed","HTST_milk"))

## PCoA 
ps.milk <- readRDS(file.path(path,"decontam/phyloseq/milk/ps.milk.rds"))
PCoA.wuf.milk <- ordinate(ps.milk, method = "PCoA", distance = "wunifrac")
sample_data(ps.milk)$SampleType <- factor(sample_data(ps.milk)$SampleType, levels = c("Raw_milk","HTST_feed","HTST_milk"))
sample_data(ps.milk)$PMA <- factor(sample_data(ps.milk)$PMA, levels = c("N","Y"))

pdf(file.path(path,"milk/PCoA_allmilk.pdf"), w = 5.3, h = 3.7)
plot_ordination(ps.milk, PCoA.wuf.milk,color="SampleType",shape = "PMA") + 
  geom_point(size=3)+ 
  scale_color_brewer(palette="Set2")+
  ggtitle("")+
  ylab("PCoA 2")+ 
  xlab("PCoA 1")+
  theme_bw(base_size = 15)
dev.off()

# Adonis testing
metadata <- as(sample_data(ps.milk), "data.frame")
# milk type
adonis2(phyloseq::distance(ps.milk, "wunifrac") ~ SampleType, data = metadata)

# Make boxplots for weighted unifrac
wuf.dist <- read.csv(file.path(path, "decontam/phyloseq/milk/PMA_MilkType_Distances_noPMA_andPMA.csv"))
wuf.dist$MilkType = factor(wuf.dist$MilkType,levels=c("Raw_milk","HTST_feed","HTST_milk"))

pdf(file.path(path, "milk/fig_SampleType/wufY.pdf"), w = 3.6, h = 3.8)
wuf.dist$facet <- "Weighted Unifrac"
ggplot(subset(wuf.dist, PMA == "Y"), aes(x=MilkType, y=weighted_distance))+
  geom_boxplot(position = position_dodge(0.77))+
  facet_wrap(~ facet) +
  xlab("")+
  theme_bw(base_size = 15) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
dev.off()

pdf(file.path(path, "milk/fig_SampleType/wufN.pdf"), w = 3.6, h = 3.8)
ggplot(subset(wuf.dist, PMA == "N"), aes(x=MilkType, y=weighted_distance))+
  geom_boxplot(position = position_dodge(0.77))+
  facet_wrap(~ facet) +
  xlab("")+
  theme_bw(base_size = 15) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
dev.off()

# ANOVA
fit <- aov(weighted_distance ~ MilkType, data = subset(wuf.dist, PMA == "Y"))
summary(fit)
# post hoc
attach(subset(wuf.dist, PMA == "Y"))
pairwise.t.test(weighted_distance, MilkType, p.adjust.method = "none")[[3]]
detach()
attach(subset(wuf.dist, PMA == "N"))
pairwise.t.test(weighted_distance, MilkType, p.adjust.method = "none")[[3]]
detach()

```

### 1.3 CheeseOutcome with 4hr (1 set) HTST feed and pasteurized milk 
```{r fig.height=3, fig.width=10}
# Define some useful functions to use 
# Define function Adonis on the beta plot
AdonisWuf <- function(ps){
  wuf.dist <- phyloseq::distance(ps, method = "wunifrac")  # dist matrix
  sampledf <- data.frame(sample_data(ps))  # data frame from the sample_data
  adonis(wuf.dist ~ CheeseOutcome, data = sampledf)
}

#merge the 4hr map with phyloseq object that has not been adjust for qPCR cell count
tree <- phy_tree(ps.milk)
tax <- tax_table(ps.milk)
otu <- otu_table(ps.milk)
ps.milk4.3 <- phyloseq(otu_table(otu, taxa_are_rows = FALSE), tax_table(tax),sample_data(samdf.4hr2), phy_tree(tree))
saveRDS(ps.milk4.3,"decontam/phyloseq/milk/ps.milk4.3.rds")

ps.milk4.3Y <- subset_samples(ps.milk4.3, PMA %in% "Y")
saveRDS(ps.milk4.3Y, "decontam/phyloseq/milk/ps.milk4.3Y.rds")
ps.milk4.3N <- subset_samples(ps.milk4.3, PMA %in% "N")
saveRDS(ps.milk4.3N, "decontam/phyloseq/milk/ps.milk4.3N.rds")

#set the order of factors
sample_data(ps.milk4.3Y)$SampleType <- factor(sample_data(ps.milk4.3Y)$SampleType, levels = c("HTST_feed","HTST_milk"))
sample_data(ps.milk4.3N)$SampleType <- factor(sample_data(ps.milk4.3N)$SampleType, levels = c("HTST_feed","HTST_milk"))

## PCoA 
# PMA treated
PCoA.wuf.milk4.3Y <- ordinate(ps.milk4.3Y, method = "PCoA", distance = "wunifrac")
# Create cluster centroid locations for only the first 3 axes
PCoA.wuf.milk4.3Y_vector <- data.frame(PCoA.wuf.milk4.3Y["vectors"])[,1:3]
# Add sample data to the vector
PCoA.wuf.milk4.3Y_vector$ID <- row.names(PCoA.wuf.milk4.3Y_vector)
PCoA.wuf.milk4.3Y_vector <- merge(PCoA.wuf.milk4.3Y_vector, 
                                sample_data(ps.milk4.3Y) %>% data.frame(),
                                by.x = "ID", by.y = "SampleID", 
                                all.x = TRUE, sort=FALSE)
#calculate the mean/centroids of the vector
PCoA.wuf.milk4.3Y_center <- ddply(PCoA.wuf.milk4.3Y_vector,
                                c("SampleType","CheeseOutcome"),
                                summarize,
                                ave1 = mean(vectors.Axis.1),
                                ave2 = mean(vectors.Axis.2),
                                ave3 = mean(vectors.Axis.3))
#merge vector and center data frame together by SampleType and CheeseOutcome columns
PCoA.wuf.milk4.3Y_merge <- merge(PCoA.wuf.milk4.3Y_vector,
                               PCoA.wuf.milk4.3Y_center,
                               by = c("SampleType","CheeseOutcome"))

#split the data 
dat_list <- split(PCoA.wuf.milk4.3Y_merge, PCoA.wuf.milk4.3Y_merge$SampleType)
#use lapply to creat plot for the 3 milk SampleType
fig_list <- lapply(dat_list, function(x){
  
  f1 <- ggplot(x, aes(vectors.Axis.1, vectors.Axis.2, color = CheeseOutcome)) + 
    geom_point(size = 1)+ 
    geom_point(aes(x = ave1, y = ave2), size = 3)+
    scale_color_manual(values=c("no_slits" = "#a6cee3", "slits" = "#1f78b4"))+
    geom_segment(aes(x = ave1, xend = vectors.Axis.1,
                     y = ave2, yend = vectors.Axis.2))+
    facet_wrap(~SampleType, scales = "fixed")+
    ylab("PCoA 2")+ 
    xlab("PCoA 1")+
    theme_bw()
})
#make the figure
legend <- get_legend(fig_list[[1]])
plot_grid(fig_list[[1]] + theme(legend.position = "none"),
          fig_list[[2]] + theme(legend.position = "none"),
          legend, nrow = 1, align="hv",  labels= c("A","B"))

# PMA untreated
PCoA.wuf.milk4.3N <- ordinate(ps.milk4.3N, method = "PCoA", distance = "wunifrac")
# Create cluster centroid locations for only the first 3 axes
PCoA.wuf.milk4.3N_vector <- data.frame(PCoA.wuf.milk4.3N["vectors"])[,1:3]
# Add sample data to the vector
PCoA.wuf.milk4.3N_vector$ID <- row.names(PCoA.wuf.milk4.3N_vector)
PCoA.wuf.milk4.3N_vector <- merge(PCoA.wuf.milk4.3N_vector, 
                                sample_data(ps.milk4.3N) %>% data.frame(),
                                by.x = "ID", by.y = "SampleID", 
                                all.x = TRUE, sort=FALSE)
#calculate the mean/centroids of the vector
PCoA.wuf.milk4.3N_center <- ddply(PCoA.wuf.milk4.3N_vector,
                                c("SampleType","CheeseOutcome"),
                                summarize,
                                ave1 = mean(vectors.Axis.1),
                                ave2 = mean(vectors.Axis.2),
                                ave3 = mean(vectors.Axis.3))
#merge vector and center data frame together by SampleType and CheeseOutcome columns
PCoA.wuf.milk4.3N_merge <- merge(PCoA.wuf.milk4.3N_vector,
                               PCoA.wuf.milk4.3N_center,
                               by = c("SampleType","CheeseOutcome"))

#split the data 
dat_list <- split(PCoA.wuf.milk4.3N_merge, PCoA.wuf.milk4.3N_merge$SampleType)
#use lapply to creat plot for the 3 milk SampleType
fig_list <- lapply(dat_list, function(x){
  
  f1 <- ggplot(x, aes(vectors.Axis.1, vectors.Axis.2, color = CheeseOutcome)) + 
    geom_point(size = 1)+ 
    geom_point(aes(x = ave1, y = ave2), size = 3)+
    scale_color_manual(values=c("no_slits" = "#a6cee3", "slits" = "#1f78b4"))+
    geom_segment(aes(x = ave1, xend = vectors.Axis.1,
                     y = ave2, yend = vectors.Axis.2))+
    facet_wrap(~SampleType, scales = "fixed")+
    ylab("PCoA 2")+ 
    xlab("PCoA 1")+
    theme_bw()
})
#make the figure
legend <- get_legend(fig_list[[1]])
plot_grid(fig_list[[1]] + theme(legend.position = "none"),
          fig_list[[2]] + theme(legend.position = "none"),
          legend, nrow = 1, align="hv",  labels= c("A","B"))

## NMDS 
# PMA treated
NMDS.wuf.milk4.3Y <- ordinate(ps.milk4.3Y, method = "NMDS", distance = "wunifrac")
# Create cluster centroid locations for only the first 3 axes
NMDS.wuf.milk4.3Y_vector <- data.frame(NMDS.wuf.milk4.3Y["points"])
# Add sample data to the vector
NMDS.wuf.milk4.3Y_vector$ID <- row.names(NMDS.wuf.milk4.3Y_vector)
NMDS.wuf.milk4.3Y_vector <- merge(NMDS.wuf.milk4.3Y_vector, 
                                sample_data(ps.milk4.3Y) %>% data.frame(),
                                by.x = "ID", by.y = "SampleID", 
                                all.x = TRUE, sort=FALSE)
#calculate the mean/centroids of the vector
NMDS.wuf.milk4.3Y_center <- ddply(NMDS.wuf.milk4.3Y_vector,
                                c("SampleType","CheeseOutcome"),
                                summarize,
                                ave1 = mean(points.MDS1),
                                ave2 = mean(points.MDS2))
#merge vector and center data frame together by SampleType and CheeseOutcome columns
NMDS.wuf.milk4.3Y_merge <- merge(NMDS.wuf.milk4.3Y_vector,
                               NMDS.wuf.milk4.3Y_center,
                               by = c("SampleType","CheeseOutcome"))

#split the data 
dat_list <- split(NMDS.wuf.milk4.3Y_merge, NMDS.wuf.milk4.3Y_merge$SampleType)
#use lapply to creat plot for the 3 milk SampleType
fig_list <- lapply(dat_list, function(x){
  
  f1 <- ggplot(x, aes(points.MDS1, points.MDS2, color = CheeseOutcome)) + 
    geom_point(size = 1)+ 
    geom_point(aes(x = ave1, y = ave2), size = 3)+
    scale_color_manual(values=c("no_slits" = "#a6cee3", "slits" = "#1f78b4"))+
    geom_segment(aes(x = ave1, xend = points.MDS1,
                     y = ave2, yend = points.MDS2))+
    facet_wrap(~SampleType, scales = "fixed")+
    ylab("NMDS 2")+ 
    xlab("NMDS 1")+
    theme_bw() +
    theme(text = element_text(size = 15)) 
})
#make the figure
#legend <- get_legend(fig_list[[1]])
pdf(file.path(path,"milk/fig/wuf_NMDSY.pdf"), w = 7.1, h = 3.2)
plot_grid(fig_list[[1]] + theme(legend.position = "none"),
          fig_list[[2]] + theme(legend.position = "none"),
          nrow = 1, align="hv",  labels= c("A","B"))
dev.off()
# Adonis test
AdonisWuf(subset_samples(ps.milk4.3Y, SampleType %in% "HTST_feed"))
AdonisWuf(subset_samples(ps.milk4.3Y, SampleType %in% "HTST_milk"))


# PMA untreated
NMDS.wuf.milk4.3N <- ordinate(ps.milk4.3N, method = "NMDS", distance = "wunifrac")
# Create cluster centroid locations for only the first 3 axes
NMDS.wuf.milk4.3N_vector <- data.frame(NMDS.wuf.milk4.3N["points"])
# Add sample data to the vector
NMDS.wuf.milk4.3N_vector$ID <- row.names(NMDS.wuf.milk4.3N_vector)
NMDS.wuf.milk4.3N_vector <- merge(NMDS.wuf.milk4.3N_vector, 
                                sample_data(ps.milk4.3N) %>% data.frame(),
                                by.x = "ID", by.y = "SampleID", 
                                all.x = TRUE, sort=FALSE)
#calculate the mean/centroids of the vector
NMDS.wuf.milk4.3N_center <- ddply(NMDS.wuf.milk4.3N_vector,
                                c("SampleType","CheeseOutcome"),
                                summarize,
                                ave1 = mean(points.MDS1),
                                ave2 = mean(points.MDS2))
#merge vector and center data frame together by SampleType and CheeseOutcome columns
NMDS.wuf.milk4.3N_merge <- merge(NMDS.wuf.milk4.3N_vector,
                               NMDS.wuf.milk4.3N_center,
                               by = c("SampleType","CheeseOutcome"))

#split the data 
dat_list <- split(NMDS.wuf.milk4.3N_merge, NMDS.wuf.milk4.3N_merge$SampleType)
#use lapply to creat plot for the 3 milk SampleType
fig_list <- lapply(dat_list, function(x){
  
  f1 <- ggplot(x, aes(points.MDS1, points.MDS2, color = CheeseOutcome)) + 
    geom_point(size = 1)+ 
    geom_point(aes(x = ave1, y = ave2), size = 3)+
    scale_color_manual(values=c("no_slits" = "#a6cee3", "slits" = "#1f78b4"))+
    geom_segment(aes(x = ave1, xend = points.MDS1,
                     y = ave2, yend = points.MDS2))+
    facet_wrap(~SampleType, scales = "fixed")+
    ylab("NMDS 2")+ 
    xlab("NMDS 1")+
    theme_bw() +
    theme(text = element_text(size = 15)) 
})
#make the figure
#legend <- get_legend(fig_list[[1]])
pdf(file.path(path,"milk/fig/wuf_NMDSN.pdf"), w = 7.1, h = 3.2)
plot_grid(fig_list[[1]] + theme(legend.position = "none"),
          fig_list[[2]] + theme(legend.position = "none"),
          nrow = 1, align="hv",  labels= c("A","B"))
dev.off()
# Adonis
AdonisWuf(subset_samples(ps.milk4.3N, SampleType %in% "HTST_feed"))
AdonisWuf(subset_samples(ps.milk4.3N, SampleType %in% "HTST_milk"))
```


Now that I have the phyloseq containing milk samples (1 set) that were collected ~4hr before the cheese was made and were not adjusted with qPCR cell counts, I just redid the alpha diversity analysis here using this phyloseq object.


```{r}
# Define ANOVA for alpha diversity testing
AlphaANOVA <- function(x){
    fit.sha <- aov(value ~ CheeseOutcome, data = subset(x, measure %in% "Observed"))
    fit.sim <- aov(value ~ CheeseOutcome, data = subset(x, measure %in% "Shannon"))
    y <- summary(fit.sha)[[1]]
    y <- rbind(y, summary(fit.sim)[[1]])
    y
}

## Alpha diversity of milk directly linked to cheese 
alpha.milk4.3Y <- estimate_richness(ps.milk4.3Y, split = TRUE, measures = alpha_measures)
alpha.milk4.3N <- estimate_richness(ps.milk4.3N, split = TRUE, measures = alpha_measures)
row.names(alpha.milk4.3Y) <- gsub("X","",row.names(alpha.milk4.3Y) )
row.names(alpha.milk4.3N) <- gsub("X","",row.names(alpha.milk4.3N) )
alpha.milk4.3Y$ID <- row.names(alpha.milk4.3Y)
alpha.milk4.3N$ID <- row.names(alpha.milk4.3N)
#add sample metadata information
alpha.milk4.3Y <- merge(alpha.milk4.3Y, data.frame(sample_data(ps.milk4.3Y)), by.x = 'ID', by.y = 'row.names')
alpha.milk4.3N <- merge(alpha.milk4.3N, data.frame(sample_data(ps.milk4.3N)), by.x = 'ID', by.y = 'row.names')
#Keep only ID variable and variables
var <- c("Observed","Chao1","Shannon","Simpson","Week","CheeseOutcome","SampleType") 
alpha.milk4.3Y <- alpha.milk4.3Y[var]
alpha.milk4.3N <- alpha.milk4.3N[var]
alpha.milk4.3Ym <- melt(alpha.milk4.3Y, id.vars = c("Week","CheeseOutcome","SampleType"),
                      variable.name = "measure",
                      value.name = "value")
alpha.milk4.3Nm <- melt(alpha.milk4.3N, id.vars = c("Week","CheeseOutcome","SampleType"),
                      variable.name = "measure",
                      value.name = "value")

# PMA treated 
ggplot(alpha.milk4.3Ym, aes(x=SampleType, y=value, fill=CheeseOutcome))+ 
  geom_boxplot() +
  scale_fill_manual(values=c("no_slits"="#a6cee3", "slits"="#1f78b4")) +
  facet_wrap(facets = ~ measure, scales = "free_y") + #strip.position="right"
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  ylab("Alpha diversity value")+
  xlab("PMA treated samples by milk type")
# ANOVA stats on alpha diversities
AlphaANOVA(subset(alpha.milk4.3Ym, SampleType %in% "HTST_feed"))

# PMA untreated 
ggplot(alpha.milk4.3Nm, aes(x=SampleType, y=value, fill=CheeseOutcome))+ 
  geom_boxplot() +
  scale_fill_manual(values=c("no_slits"="#a6cee3", "slits"="#1f78b4")) +
  facet_wrap(facets = ~ measure, scales = "free_y") + #strip.position="right"
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  ylab("Alpha diversity value")+
  xlab("PMA untreated samples by milk type")
AlphaANOVA(subset(alpha.milk4.3Nm, SampleType %in% "HTST_feed"))


## Bacterial cell count of milk directly linked to cheese 
samdf4.3Y <- sample_data(ps.milk4.3Y) %>% data.frame()
samdf4.3N <- sample_data(ps.milk4.3N) %>% data.frame()
qPCR4.3Y <- merge(x = qPCR, y = samdf4.3Y, by.x = "SampleID", by.y = "SampleID")
qPCR4.3N <- merge(x = qPCR, y = samdf4.3N, by.x = "SampleID", by.y = "SampleID")
#Keep only ID variable and variables
qPCR4.3Y <- qPCR4.3Y[c("SampleType","CheeseOutcome.y","CollectionTime","Week.y","LogCellsML")]
qPCR4.3N <- qPCR4.3N[c("SampleType","CheeseOutcome.y","CollectionTime","Week.y","LogCellsML")]
qPCR4.3Ym <- melt(qPCR4.3Y, id.vars = c("Week.y","CheeseOutcome.y","SampleType","LogCellsML"),
                  variable.name = "measure",
                  value.name = "value")
qPCR4.3Nm <- melt(qPCR4.3N, id.vars = c("Week.y","CheeseOutcome.y","SampleType","LogCellsML"),
                  variable.name = "measure",
                  value.name = "value")
```

```{r fig.height=3, fig.width=4.3}
# PMA treated 
qPCR4.3Ym$facet <- "Log cells/ml"
pdf(file.path(path,"milk/alpha/log4.3Y.pdf"), w = 5.5, h = 3.3)
ggplot(qPCR4.3Ym, aes(x=SampleType, y=LogCellsML, fill=CheeseOutcome.y))+ 
  geom_boxplot() +
  scale_fill_manual(values=c("no_slits"="#a6cee3", "slits"="#1f78b4")) +
  theme_bw(base_size = 15)+
  facet_wrap(~ facet)+
  ylab("Log cells/ml")
dev.off()
aov(LogCellsML ~ CheeseOutcome.y, 
    data = subset(qPCR4.3Ym, SampleType %in% "HTST_feed")) %>% summary()
aov(LogCellsML ~ CheeseOutcome.y, 
    data = subset(qPCR4.3Ym, SampleType %in% "HTST_milk")) %>% summary()

# PMA untreated
qPCR4.3Nm$facet <- "Log cells/ml"
pdf(file.path(path,"milk/alpha/log4.3N.pdf"), w = 5.5, h = 3.3)
ggplot(qPCR4.3Nm, aes(x=SampleType, y=LogCellsML, fill=CheeseOutcome.y))+ 
  geom_boxplot() +
  scale_fill_manual(values=c("no_slits"="#a6cee3", "slits"="#1f78b4")) +
  theme_bw(base_size = 15)+
  facet_wrap(~ facet)+
  ylab("Log cells/ml")
dev.off()
```

## Import the qPCR results for Turicibacter counts in cheese and milk
```{r fig.height=3, fig.width=4.3}
qPCR.Tur <- read.csv("mapping/Good_data_Turicibacter_qPCR_NoLowValue.csv")
# subset by PMA treated and untreated
qPCR.TurY <- qPCR.Tur[qPCR.Tur$PMA == "Y",]
qPCR.TurN <- qPCR.Tur[qPCR.Tur$PMA == "N",] 

# melt dataframe to be suited for ploting
qPCR.TurYm <- melt(qPCR.TurY, id.vars = c("Week","CheeseOutcome","SampleType","LOG"),
                  variable.name = "measure",
                  value.name = "value")
qPCR.TurNm <- melt(qPCR.TurY, id.vars = c("Week","CheeseOutcome","SampleType","LOG"),
                  variable.name = "measure",
                  value.name = "value")

# Set levels to the SampleType
qPCR.TurYm$SampleType <- factor(qPCR.TurYm$SampleType, levels = c("Cheese_0D","Cheese_30D","Cheese_90D","Cheese_120D","HTST_feed","HTST_milk"))
qPCR.TurNm$SampleType <- factor(qPCR.TurNm$SampleType, levels = c("Cheese_0D","Cheese_30D","Cheese_90D","Cheese_120D","HTST_feed","HTST_milk"))


# Make similar figure as the total cell count by creating a slit.plot function
## SHOULD HAVE DONE THIS A LONG TIME AGO, WOULD BE MUCH SMARTER
slit.plot <- function(x){
  ggplot(x, aes(x=SampleType, y=LOG, fill=CheeseOutcome))+ 
  geom_boxplot() +
  scale_fill_manual(values=c("no_slits"="#a6cee3", "slits"="#1f78b4")) +
  theme_bw()+
  ylab("Turicibacter")
}
# ANOVA tests for taxa 
## I didn't use Wald test because I couldn't get it to work to establish a lm or other model
ChouANOVA <- function(x){
  fit <- aov(LOG ~ CheeseOutcome, data = x)
  summary(fit)
}


slit.plot(qPCR.TurYm[qPCR.TurYm$SampleType %in% c("Cheese_0D","Cheese_30D","Cheese_90D","Cheese_120D"),]) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  xlab("PMA treated samples by cheese type")
ChouANOVA(qPCR.TurYm[qPCR.TurYm$SampleType %in% c("Cheese_0D"),])
ChouANOVA(qPCR.TurYm[qPCR.TurYm$SampleType %in% c("Cheese_30D"),])
ChouANOVA(qPCR.TurYm[qPCR.TurYm$SampleType %in% c("Cheese_90D"),])
ChouANOVA(qPCR.TurYm[qPCR.TurYm$SampleType %in% c("Cheese_120D"),])

slit.plot(qPCR.TurNm[qPCR.TurNm$SampleType %in% c("Cheese_0D","Cheese_30D","Cheese_90D","Cheese_120D"),]) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  xlab("PMA untreated samples by cheese type")
ChouANOVA(qPCR.TurNm[qPCR.TurNm$SampleType %in% c("Cheese_0D"),])
ChouANOVA(qPCR.TurNm[qPCR.TurNm$SampleType %in% c("Cheese_30D"),])
ChouANOVA(qPCR.TurNm[qPCR.TurNm$SampleType %in% c("Cheese_90D"),])
ChouANOVA(qPCR.TurNm[qPCR.TurNm$SampleType %in% c("Cheese_120D"),])


slit.plot(qPCR.TurYm[qPCR.TurYm$SampleType %in% c("HTST_feed","HTST_milk"),]) +
  xlab("PMA treated samples by milk type")
ChouANOVA(qPCR.TurNm[qPCR.TurNm$SampleType %in% c("HTST_feed"),])
ChouANOVA(qPCR.TurNm[qPCR.TurNm$SampleType %in% c("HTST_milk"),])
  
slit.plot(qPCR.TurNm[qPCR.TurNm$SampleType %in% c("HTST_feed","HTST_milk"),]) +
  xlab("PMA treated samples by milk type")
```

