---
title: "pilot_cheese_inoculation_sequencing"
author: "Zeya Zhengyao Xue"
date: "2/6/2019"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.path = "figs/")
```

This is the file to analysis pilot cheese project, including 
1. pilot cheeses （high level isolate inoculation）
2. milk and cheeese where the isolates are from

## Setting up packages and working directory 
```{r}
library(phyloseq);packageVersion("phyloseq")
library(tidyr)
library(ggplot2)
library(reshape2)
library(plyr)
library(dplyr)
library(PMCMR)
library(vegan)
library(RColorBrewer)
library(superheat)
library(DESeq2);packageVersion("DESeq2")

path <- "~/Google Drive File Stream/My Drive/UC_Davis/Marco_lab/milk_microbiota/2017_Hilmar_sampling/Sequencing/QIIME2/"
```


## Import QIIME2 output and make phyloseq object
```{r}
# Transpose the QIIME2 output in excel
SeqTab <- read.table(file.path(path,"Jan/feature-table-mc2.txt"), header = TRUE, stringsAsFactors = FALSE)
colnames(SeqTab) <- gsub("X","",colnames(SeqTab))
row.names(SeqTab) <- SeqTab$OTUID
SeqTab <- SeqTab[,-1]
SeqTab <- as.matrix.data.frame(SeqTab)

# Parse out taxonomic assignment in excel and remove confidence column
TaxTab <- read.csv(file.path(path,"Jan/taxonomy-mc2.csv"))
# remove species column
TaxTab <- subset(TaxTab, select = -c(Species))
# organized as a matrix for phyloseq import
rownames(TaxTab) <- TaxTab$Feature.ID
TaxTab <- TaxTab[,-1]
TaxTab <- as.matrix.data.frame(TaxTab)

# Read in metadata file
# In the samdf file, L.fermentum R2 was removed due to potential contamination from 
samdf <- read.csv(file.path(path,"Jan/samdf_Jan_2019.csv"))
rownames(samdf) <- samdf$SampleID

# Read in exported rooted tree 
tree <- read_tree(file.path(path, "Jan/tree.nwk"))

### Merge as one phyloseq object
ps <- phyloseq(otu_table(SeqTab,taxa_are_rows = TRUE), tax_table(TaxTab), sample_data(samdf), phy_tree(tree))
ps # 282 taxa and 29 samples



```

## Filter ASVs based on taxonomy and abundance 
```{r}
# Show available ranks in the dataset
rank_names(ps)
# Create table, number of features for each phylum
table(tax_table(ps)[, "Phylum"], exclude = NULL)
# remove asv that was not assigned at phylum level
ps.Phy <- subset_taxa(ps, !is.na(Phylum) & !Phylum %in% c("","None"))
ps.Phy #  279 taxa and 29 samples 
# Explore abundance to decide filter threshold
## Compute prevalence of each feature, store as data.frame
prevdf <- apply(X = otu_table(ps.Phy),
                MARGIN = ifelse(taxa_are_rows(ps.Phy), yes = 1, no = 2),
                FUN = function(x)(sum(x>0)))
# Add taxonomy and total read counts to this data.frame
prevdf <- data.frame(Prevalence = prevdf,
                     TotalAbundance = taxa_sums(ps.Phy),
                     tax_table(ps.Phy))
# plot the abudance per phylum 
ggplot(prevdf, aes(TotalAbundance, Prevalence / nsamples(ps.Phy),color=Phylum)) +
  # Include a guess for parameter
  geom_hline(yintercept = 0.01,  linetype = 2) + geom_point(size = 2, alpha = 0.7) +
  scale_x_log10() +  xlab("Total Abundance") + ylab("Prevalence [Frac. Samples]") +
  facet_wrap(~Phylum) + theme(legend.position="none")

# Remove very low abudance ASVs independent of sample prevalence
x = taxa_sums(ps.Phy)
keepTaxa = which((x / sum(x)) > 0.00005)
ps.Phy05 <- prune_taxa(names(keepTaxa), ps.Phy)
ps.Phy05 # 175 taxa and 29 samples
```

## Determine rarefaction level for Alpha and Beta analysis later on
```{r}
set.seed(123)
# Read in function to calculate alpha diversity and rarefaction levels
calculate_rarefaction_curves <- dget(file.path("~/src/Marco_Lab/FUN_calculate_rarefaction.R"))
rarefaction_curve_data <- calculate_rarefaction_curves(ps.Phy, c("Observed", "Shannon","Simpson","Chao1"), 
                                                       rep(c(1, 1000, 2500, 4000, 6000, 10000, 20000), each = 10))
rarefaction_curve_data$SampleID <- gsub("X","",rarefaction_curve_data$SampleID)
# summarize alpha diversity
rarefaction_curve_data_summary <- ddply(rarefaction_curve_data, c('Depth', 'SampleID', 'measures'),
                                        summarise, 
                                        Alpha_diversity_mean = mean(Alpha_diversity), 
                                        Alpha_diversity_sd = sd(Alpha_diversity))
# Add sample data
rarefaction_curve_data_summary_verbose <- merge(rarefaction_curve_data_summary, 
                                                data.frame(sample_data(ps.Phy)), 
                                                by.x = 'SampleID', by.y = 'row.names')
rarefaction_curve_data_summary_verbose <- rarefaction_curve_data_summary_verbose[, -1]
# plot out rarefaction curve
ggplot(rarefaction_curve_data_summary_verbose, aes(x = Depth, y = Alpha_diversity_mean, 
                                                   ymin = Alpha_diversity_mean - Alpha_diversity_sd, 
                                                   ymax = Alpha_diversity_mean + Alpha_diversity_sd,
                                                   color = SampleID, group = SampleID)) +
  scale_fill_brewer(palette = "Set2") + 
  geom_line()+
  geom_pointrange(size=0.1)+
  facet_wrap(facets = ~ measures, scales = "free")
# export the library size to a csv file
write.csv(sample_sums(ps.Phy05), file.path(path, "Jan/library_size_ps.Phy05.csv"))

# The diversity leveled out after 2000 seqs per sample, rarefy at 6100 reads
# to keep the sample with the fewest reads
```


## Divide the ps object for milk and cheese
### For taxonomy analysis based on "ps.REC.glom"
The goal is to 
(1) Confirm that the isolates do exist in the milk that they were isolated from 
(2) The pilot cheeses made with the isolate had similar slit profile 
(3) the milk containing the isolates had similar slits profile compared to study 1
[but I don't have no slits samples from the same cheese study 2]
```{r}
# Clean up the taxonomy level
#make the deepest taxonomic identification in TaxTab as the recommened taxonomy (REC)
RECps <- dget("~/src/Marco_Lab/FUN_RECps.R") # Read in RECps function
ps.Phy.REC <- RECps(ps.Phy05@tax_table, ps.Phy05)
ps.Phy.REC # 175 taxa and 29 samples
ps.REC.glom <- tax_glom(ps.Phy.REC, taxrank = "REC", NArm = FALSE)
ps.REC.glom #  85 taxa and 29 samples 

# Subset to contain milk samples 
ps.milk <- subset_samples(ps.REC.glom, SampleType %in% c("HTST_feed", "HTST_milk"))
ps.milk #  85 taxa and 18 samples 
# Merge with ps.milk from cheese study 1 as well


# Subset to contain cheese samples ()
ps.cheese <- subset_samples(ps.REC.glom, SampleType %in% "Pilot_cheese_5D")
ps.cheese #  85 taxa and 11 samples 
ps.cheese.noL <- subset_taxa(ps.cheese, !Genus %in% "Lactococcus")
ps.cheese.noL # 84 taxa and 11 samples

# Define function to make Bar plots 
#### Bar plot #####
SlitBar <- function(ps, var1, var2, taxa, lvl1, lvl2, mycol, path.out, w, h){  
  ps <- ps %>% transform_sample_counts(function(x) x/sum(x) ) 
  
  TaxTab <- tax_table(ps) %>% as.data.frame()
  taxa_names(ps) <- TaxTab$REC
  allTaxa <- taxa_names(ps)
  ps.other <- prune_taxa(allTaxa[!(allTaxa %in% taxa)], ps)
  taxa.other <- ps.other@tax_table[,7] %>% as.character() # 7 for REC level  
  
  # change the taxa that are not the expected taxa to "Other" 
  TaxTab2 <- psmelt(ps)
  # merge/average per Sample and CheeseOutcome
  TaxTab2_agg <- ddply(TaxTab2, c(var1, var2, "REC"),
                       summarise,
                       mean = mean(Abundance))
  colnames(TaxTab2_agg)[1] <- var1
  colnames(TaxTab2_agg)[2] <- var2
  TaxTab2_agg$REC <- as.character(TaxTab2_agg$REC)
  TaxTab2_agg[TaxTab2_agg$REC %in% taxa.other,]$REC <- "Other"
  
  # Set orders for taxonomy and sample
  TaxTab2_agg$REC <- factor(TaxTab2_agg$REC, levels = taxa)
  TaxTab2_agg[[var1]] <- factor(TaxTab2_agg[[var1]], levels = lvl1)
  TaxTab2_agg[[var2]] <- factor(TaxTab2_agg[[var2]], levels = lvl2)
  
  pdf(path.out, w, h)
  p <- ggplot(TaxTab2_agg, aes(x = get(var1), y = mean, fill = REC)) +
    geom_bar(stat = "identity") + 
    scale_fill_manual(values = mycol)+
    theme_bw(base_size = 20) +
    facet_grid(~ get(var2), scales = "free_x")+
    # Remove x axis title
    theme(axis.title.x = element_blank(), axis.text.x = element_text(angle = 0, hjust = 1),
          text = element_text(size = 16)) + 
    guides(fill = guide_legend(reverse = FALSE, keywidth = 1, keyheight = 1)) +
    ylab("Relative abundance")
  
  # convert ggplot object to grob object
  gp <- ggplotGrob(p)
  # optional: take a look at the grob object's layout 
  # gtable::gtable_show_layout(gp)
  # get gtable columns corresponding to the facets 
  facet.columns <- gp$layout$l[grepl("panel", gp$layout$name)]
  # get the number of unique x-axis values per facet (1 & 3, in this case)
  x.var <- sapply(ggplot_build(p)$layout$panel_scales_x, 
                  function(l) length(l$range$range))
  # change the relative widths of the facet columns based on how many unique 
  # x-axis values are in each facet
  gp$widths[facet.columns] <- gp$widths[facet.columns] * x.var
  # plot result
  grid::grid.draw(gp)
  dev.off()
}

comm17 <- c("Acinetobacter", "Bacillus", "Bacillaceae", "Clostridium", 
            "Clostridiaceae", "Clostridiales", "Enterobacteriaceae", "Lactococcus", 
            "Lactobacillus", "Lactobacillaceae","Leuconostoc", "Pseudomonas",
            "Staphylococcus", "Streptococcus", "Thermus", "Turicibacter", "Other")
col17 <- c("#A6CEE3", "#438EC0", "#1f78b4","#b3e2cd", 
           "#98D277", "#3BA432", "#B89B74", "#E62F27", 
           "#F16667", "#fbb4ae", "#F9A963", "#ff7f00", 
           "#C3AAD2", "#7D54A5", "#B9A499", "#EAD27A", "#B15928")
SlitBar(ps.milk, var1 = "CollectionGroup", var2 = "SampleType",
        lvl1 = c("Beginning","Middle","End"),
        lvl2 = c("HTST_feed","HTST_milk"),
        taxa = comm17, mycol = col17, 
        path.out = file.path(path, "figs/milk_CollectionGroup_SampleType.pdf"), 6.8, 5.5)
SlitBar(ps.milk, var2 = "SampleType", var1 = "PMA.treatment",
        lvl2 = c("HTST_feed","HTST_milk"),
        lvl1 = c("N"),
        taxa = comm17, mycol = col17, 
        path.out = file.path(path, "figs/milk_SampleType.pdf"), 4.8, 5.5)
#SlitBar(ps.cheese, var1 = "Bacterial_isolate", var2 = "SampleType",
#        taxa = comm16, mycol = col16,
#        path.out = file.path(path, "figs/cheese_Bacterial_isolate_SampleType.pdf"), 6, 9)
SlitBar(ps.cheese, var1 = "Species", var2 = "SampleType",
        lvl1 = c("Saline","Lactobacillus plantarum","Lactobacillus fermentum ",
                 "Leuconostoc lactis ","Leuconostoc mesenteroides "),
        lvl2 = c("Pilot_cheese_5D"),
        taxa = comm17, mycol = col17,
        path.out = file.path(path, "figs/cheese_Species_SampleType.pdf"), 6, 5.5)

# comm16 doesn't have Lactococcus 
comm16 <- c("Acinetobacter", "Bacillus", "Bacillaceae", "Clostridium", 
            "Clostridiaceae", "Clostridiales", "Enterobacteriaceae", 
            "Lactobacillus", "Lactobacillaceae","Leuconostoc", "Pseudomonas",
            "Staphylococcus", "Streptococcus", "Thermus", "Turicibacter", "Other")
col16 <- c("#A6CEE3", "#438EC0", "#1f78b4","#b3e2cd", 
           "#98D277", "#3BA432", "#B89B74", 
           "#F16667", "#fbb4ae", "#F9A963", "#ff7f00", 
           "#C3AAD2", "#7D54A5", "#B9A499", "#EAD27A", "#B15928")
```


# DESeq2 
```{r}
# calculate geometric means prior to estimate size factors
gmMean = function(x, na.rm=TRUE){
  exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
}
# Cheese Outcome Differential abundance and generate diagnostic plots and write sig tables 
## fold change is 1. sample groups compare with saline 2. sample vs Lp
## Uses default Benjamini-Hochberg pvalue adjust
HighIsoCheeseDA <- function(ps) {
  sample_data(ps)$Species <- factor(sample_data(ps)$Species)
  psdds <- phyloseq_to_deseq2(ps, design = ~ Species)
  
  geoMeans <- apply(counts(psdds), 1, gmMean)
  psdds <- estimateSizeFactors(psdds, geoMeans = geoMeans)
  dds <- DESeq(psdds, test = "Wald", fitType = "local") 
  plotDispEsts(dds, ylim = c(1e-6, 1e2)) 
  alpha <- 0.1 #  padj for indepenent filtering and expect FDR < alpha
  
  resSLf <- results(dds, contrast = c("Species", "Saline", "Lactobacillus fermentum "),
                    alpha = alpha, lfcThreshold = 0, altHypothesis = "greaterAbs")
  resSLl <- results(dds, contrast = c("Species", "Saline", "Leuconostoc lactis "),
                    alpha = alpha, lfcThreshold = 0, altHypothesis = "greaterAbs")
  resSLm <- results(dds, contrast = c("Species", "Saline", "Leuconostoc mesenteroides "),
                    alpha = alpha, lfcThreshold = 0, altHypothesis = "greaterAbs")
  resLpLf <- results(dds, contrast = c("Species", "Lactobacillus plantarum", 
                                       "Lactobacillus fermentum "),
                     alpha = alpha, lfcThreshold = 0, altHypothesis = "greaterAbs")
  resLpLl <- results(dds, contrast = c("Species", "Lactobacillus plantarum", 
                                       "Leuconostoc lactis "),
                     alpha = alpha, lfcThreshold = 0, altHypothesis = "greaterAbs")
  resLpLm <- results(dds, contrast = c("Species", "Lactobacillus plantarum", 
                                       "Leuconostoc mesenteroides "),
                     alpha = alpha, lfcThreshold = 0, altHypothesis = "greaterAbs")  
 
  # Define function to write out csv tables 
  f <- function(res, path.out){
    mcols(res, use.names = TRUE)
    sigtab <- res
    if (nrow(sigtab) > 0){
      sigtab <- cbind(as(sigtab, "data.frame"), 
                      as(tax_table(ps)[rownames(sigtab), ], "matrix"))
      write.csv(sigtab, file.path(path.out))
    }
  }
  
  f(resSLf, file.path(path, "DESeqTab/High_resSLf.csv"))
  f(resSLl, file.path(path, "DESeqTab/High_resSLl.csv"))
  f(resSLm, file.path(path, "DESeqTab/High_resSLm.csv"))
  f(resLpLf, file.path(path, "DESeqTab/High_resLpLf.csv"))
  f(resLpLl, file.path(path, "DESeqTab/High_resLpLl.csv"))
  f(resLpLm, file.path(path, "DESeqTab/High_resLpLm.csv"))
}

HighIsoCheeseDA(ps.cheese)
```


## Prportion analysis of isolates
```{r}
ps.iso <- subset_taxa(ps.milk %>% transform_sample_counts(function(x) x/sum(x)),  
                      REC %in% c("Lactobacillus", "Lactobacillaceae", "Leuconostoc"))
ps.iso # 3 taxa and 18 samples

IsoBar <- function(ps, var1, var2, lvl1, lvl2, mycol, path.out, w, h){
  TaxTab.iso <- psmelt(ps) 
  TaxTab.iso_agg <- ddply(TaxTab.iso, c(var1, var2, "REC"), 
                           summarise,
                           N = length(Abundance),
                           mean = mean(Abundance),
                           sd = sd(Abundance),
                           se = sd / sqrt(N))
  
  # merge/average per var1 and var2
  colnames(TaxTab.iso_agg)[1] <- var1
  colnames(TaxTab.iso_agg)[2] <- var2
  
  # Set orders for taxonomy and sample
  TaxTab.iso_agg$REC <- factor(TaxTab.iso_agg$REC, levels = c("Lactobacillus", 
                                                              "Lactobacillaceae",
                                                              "Leuconostoc"))
  TaxTab.iso_agg[[var1]] <- factor(TaxTab.iso_agg[[var1]], levels = lvl1)
  TaxTab.iso_agg[[var2]] <- factor(TaxTab.iso_agg[[var2]], levels = lvl2)
  
  pdf(path.out, w, h)
  print(ggplot(TaxTab.iso_agg, aes(x = get(var1), y = mean, fill = REC)) +
      geom_bar(stat = "identity", position = "dodge") + 
      geom_errorbar(aes(ymin = mean-se, ymax = mean+se), 
                    width = 0.8, position = position_dodge(.9)) +
      scale_fill_manual(values = mycol)+
      theme_bw(base_size = 20) +
      facet_grid(~ get(var2), scales = "free_x")+
      # Remove x axis title
      theme(axis.title.x = element_blank(), 
            axis.text.x = element_text(angle = 0, hjust = 1), 
            text = element_text(size = 16)) + 
      guides(fill = guide_legend(reverse = FALSE, keywidth = 1, keyheight = 1)) +
      ylab("Relative abundance"))
  dev.off()
}

IsoBar(ps.iso, var2 = "SampleType", var1 = "Species", 
       mycol = c("#F16667", "#fbb4ae", "#F9A963"),
       lvl2 = c("HTST_feed","HTST_milk"), 
       lvl1 = c("Lactobacillus fermentum ",
                "Leuconostoc lactis ",
                "Leuconostoc mesenteroides "), 
       path.out = file.path(path, "figs/Isolates_in_milk.pdf"), 6, 5.5)

# Box plot 
## This is the figure I decided to use in paper 
TaxTab.iso <- psmelt(subset_samples(ps.iso, SampleID %in% c("3490","3498","3502",
                                                            "3504","3512","3518",
                                                            "3850","3860","3872")))
                                    # subset to contain directly linked milk samples with isolates
# Subset to contain taxa that were directly linked with isolates in milk
index.Lb <- which(with(TaxTab.iso, Species %in% "Lactobacillus fermentum " & 
                         REC %in% c("Lactobacillus","Lactobacillaceae") & 
                         SampleType %in% "HTST_milk"))
index.Lm <- which(with(TaxTab.iso, Species %in% "Leuconostoc mesenteroides " & 
                         REC %in% "Leuconostoc" &  SampleType %in% "HTST_feed"))
index.Ll <- which(with(TaxTab.iso, Species %in% "Leuconostoc lactis " & 
                         REC %in% "Leuconostoc" &  SampleType %in% "HTST_feed"))
TaxTab.iso2 <- TaxTab.iso[c(index.Lb, index.Lm, index.Ll),]

# Set orders for taxonomy and sample
TaxTab.iso2$REC <- factor(TaxTab.iso2$REC, levels = c("Lactobacillus", 
                                                      "Lactobacillaceae",
                                                      "Leuconostoc"))
pdf(file.path(path, "figs/Isolates_in_milk_box.pdf"), 6, 5.5)
ggplot(TaxTab.iso2, aes(x = Species, y = Abundance, fill = REC)) +
  geom_boxplot() + 
  scale_fill_manual(values = c("Lactobacillus" = "#fbb4ae",
                               "Lactobacillaceae" = "#F16667",
                               "Leuconostoc" = "#F9A963"))+
  theme_bw(base_size = 20) +
  facet_grid(~ SampleType, scales = "free_x")+
  # Remove x axis title
  theme(axis.title.x = element_blank(), 
        axis.text.x = element_text(angle = 0, hjust = 1), 
        text = element_text(size = 16)) + 
  guides(fill = guide_legend(reverse = FALSE, keywidth = 1, keyheight = 1)) +
  ylab("Relative abundance")
dev.off()
```



## Chemical analysis portion 
```{r}
chem <- read.csv(file.path(path,"Sample_chemical_analysis_subset.csv"))
chemm <- melt(chem, id.vars = c("Marco_lab_sample_label",
                                "Collection_group", "Cheese_D90_quality", 
                                "Cheese_D90_slits", "Sample_type"),
              na.rm = TRUE)


pdf(file.path(path, "figs/Cheese90D_chem.pdf"), w = 6, h = 5.5)
ggboxplot(subset(chemm, Sample_type == "Cheese_day90"), 
          x = "Cheese_D90_slits", y = "value",
          facet.by = "variable", scales = "free_y", 
          fill = "Cheese_D90_slits", palette = c("#a6cee3","#1f78b4")) + 
  stat_compare_means(method = "t.test", label =  "p.signif", hide.ns = TRUE)
dev.off()

pdf(file.path(path, "figs/Cheese5D_chem.pdf"), w = 6, h = 5.5)
ggboxplot(subset(chemm, Sample_type == "Cheese_day5"), 
          x = "Cheese_D90_slits", y = "value",
          facet.by = "variable", scales = "free_y", 
          fill = "Cheese_D90_slits", palette = c("#a6cee3","#1f78b4")) + 
  stat_compare_means(method = "t.test", label =  "p.signif", hide.ns = TRUE)
dev.off()
```

